------------------------------------------------------------------------

---
title: "manuscript_full"
header-includes:
 - \usepackage{mathtools,amssymb,amsmath,amsthm}
 - \newtheorem{assumption}{Assumption}
output:
  html_document: default
  pdf_document: default
editor: visual
params:
  show: no
date: "2024-06-03"
---

```{r setup, include = FALSE, echo = FALSE, warning = FALSE, message = FALSE, dev = "CairoPNG"}

options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
	echo = FALSE,
	fig.topcaption = FALSE,
	message = FALSE,
	dev = "CairoPNG",
	dpi = 600,
	results = "hide"
)
  # adjust zoom here
  # screenshot.opts = list(zoom = 5))

# loadin libraries
library(ratioPCR)
# library(qpcR)
library(ggplot2)
# library(ggbreak)
# library(ggpubr)
# library(ggsci)
# library(dplyr)
# library(reshape2)
# library(RColorBrewer)
# library(ggh4x) # ggh4x::strip_themed
# library(cowplot)
# library(latex2exp)
# library(grid)
# library(ggthemes)
# library(viridis)
# library(devEMF)
# library(lemon)
suppressPackageStartupMessages(library(purrr))
ppath <- "/scicore/home/sinnreic/kinter/Rprojects.old/rPCR/"

fpath <- paste0(ppath,"manuscript/NAR/figures/")

calibrationFile <- "documents/extdata/20191120AS calibration.xlsx"


# importing data
import <- rpcr_wrap(rpcr_readTF(paste0(ppath,calibrationFile)))
import <- import[!import$sample == "neg. ctrl", ]
import <- rpcr_calibDetect(import,qual="")

# variable definiton
genelist <- c("NUMA1","SYNE1","KIF13A")
datalist <- c("F","bF")
modellist <- c("linear","single")
import <- rpcr_calibDetect(import, qual="")
# import$sample <- factor(import$sample)
 import$USI <- import$FA/(import$FA+import$FB)
 import$bUSI <- import$bFA/(import$bFA+import$bFB)

# List of standards used for the calibration curve
psiCal <- c(0, 0.05, 0.15, 0.25, 0.35, 0.5, 0.65, 0.75, 0.85, 0.95, 1)
tPSI <- rep(psiCal, each = 3)

rpcr_palette <- c("#E64B35FF","#4DBBD5FF","#00A087FF","#3C5488FF","#F39B7FFF","#8491B4FF","#91D1C2FF","#DC0000FF","#7E6148FF","#B09C85FF")


rpcr_col_target <- rpcr_palette[-c(1,3)]
rpcr_col_probe <- c(rgb(199, 235, 211, maxColorValue = 255),rgb(255, 255, 190, maxColorValue = 255),rgb(255, 190, 190, maxColorValue = 255))
rpcr_col_probe <- c("#E64B35FF",rgb(255, 255, 100, maxColorValue = 255),"#00A087FF")
rpcr_col_sample <- colorRampPalette(c("grey90", "grey30"))  

colplot = TRUE


if (!exists("figures")) figures <- list()
if (!exists("res")) res <- list()


standardSettings <- data.frame(signal = "Rn",
                               tres = "0:0",
                               rtype = "dynamic:dynamic",
                               roff = "0:0",
                               model = "linear",
                               method = "incl:total",
                               calib = "1:0:0:1",
                               average = TRUE,
                               offset=1,
                               derivative = FALSE,
                               baseline = FALSE)

psave <- function(filename, plot, device="pdf",path=fpath,units="cm",dpi=300, ...)
  ggsave(filename=paste0(filename,".",device),plot=plot,device=device, path=path,units=units,dpi=dpi, ...)

```

```{r}

# does not work
```

```{r}
# {r Figure_1, fig.width = 14, fig.height = 10, fig.dpi = 600}


analysis <- "Overview PCR"

colplot <- TRUE
target_color <- rpcr_col_target[1:3]
iso_color <- if (colplot) rpcr_col_probe[c(1, 3)] else rev(rpcr_col_sample(2))



plots <- rpcr_plotCalibSummary(import, colplot = colplot, color_probe = rpcr_col_probe)


df <- tidyr::pivot_longer(import,
   cols = c("FA","FB","bFA","bFB"),   # columns to reshape
    names_to = "variable",       # new column for former column names
    values_to = "value"          # new column for values
  )


df <- df[df$sample %in% c("0", "100"), ]

strip <-  ggh4x::strip_themed(background_y = if (colplot) ggh4x::elem_list_rect(fill = target_color) else element_blank(),background_x = element_blank())

relabel <- as_labeller(c("bFA" = "inclusion", "bFB" = "exclusion","0" = "0%", "100" = "100%.","KIF13A" = "KIF13A","NUMA1" = "NUMA1","SYNE1" = "SYNE1","template" = "template isoform","cross"="cross-\nreactive","correct"="correct\nisoform"))


df$info <- "template"
df <- df[df$cycle==45&df$variable %in% c("bFA","bFB"),]
df$cat <- FALSE
df[df$sample==0,]$cat <- df[df$sample==0,]$variable=="bFB"
df[df$sample==100,]$cat <- df[df$sample==100,]$variable=="bFA"

df$cat <- c("correct","cross")[2-1*df$cat]

p3 <- ggplot() + stat_summary(data=df,aes(x=sample,y=value,group=variable), fun = mean, position="dodge",geom="col", size=0.5,lty=1,col="black",fill=rep(iso_color,6),width = 0.7) + # ,fill=rep(iso_color,6)) +
   # scale_fill_manual(values=iso_color) + 
  stat_summary(data=df,aes(x=sample,y=value),fun.data = mean_se, geom = "errorbar", 
     width = 0.5, size = 0.5, col = "black") + 
  scale_y_continuous(breaks = c(0,5,10),labels = scales::number_format(accuracy = 1)) +
  ggh4x::facet_grid2(target ~ cat,strip=strip,labeller = relabel,scales="free_x", independent = "x") +
  theme_rpcr()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme( axis.text.y=element_text(size = 10),panel.spacing = unit(1, "lines"),strip.text.y = element_blank(),strip.text.x=element_text(size = 10,face = "plain")) +
  scale_x_discrete(labels=c("0%", "100%")) +
  labs(x ="PSI of samples \n ", y = "Flourescence signal [deltaRn]")+
            theme(axis.text = element_text(size = 8),axis.title = element_text(face="plain"),axis.text.y = element_text(face="plain"))

  


figures[[analysis]]$A <- ggpubr::ggarrange(plots[[1]][[3]],
                plots[[1]][[4]] + 
                theme(strip.background.y = element_blank(),strip.text.y = element_blank()) + 
                labs(title = " ",x ="Fluorescence signal [deltaRn]\n(inclusion probe)", y = "Fluorescence signal [deltaRn]\n(exclusion probe)"),
                p3, ncol = 3, nrow = 1,labels = c("A","B","C"),common.legend = TRUE, legend="bottom", widths = c(4, 2,2))

figures[[analysis]]$B <- ggpubr::ggarrange(plots[[2]][[3]],
                plots[[2]][[4]]+ 
                theme(strip.background.y = element_blank(),strip.text.y = element_blank()) + 
                labs(title = " ",x ="Fluorescence signal [deltaRn]\n(inclusion probe)", y = "Fluorescence signal [deltaRn]\n(exclusion probe)"),
                p3, ncol = 3, nrow = 1,labels = c("A","B","C"),common.legend = TRUE, legend="bottom",widths = c(4, 2,2))



figures[[analysis]]

ppp1 <- rpcr_plotCalibSummary(import, colplot = FALSE, color_probe = rpcr_col_probe)[[2]][[1]]
ppp2 <- rpcr_plotCalibSummary(import, colplot = FALSE, color_probe = rpcr_col_probe)[[1]][[1]]

ggsave(paste0(fpath,"Cross1.png"), p3,width=4, dpi=300,units="cm")
ggsave(paste0(fpath,"Figure2_panelc.png"),plots[[2]][[3]],width=6,height=6,dpi=300,units="cm")
ggsave(paste0(fpath,"Figure2_paneld.png"),ppp1,width=8,height=6,dpi=300,units="cm")
ggsave(paste0(fpath,"Figure2_panele.png"),ppp2,width=8,height=6,dpi=300,units="cm")
ggsave(paste0(fpath,"Figure1_A.pdf"),figures[[1]]$A,width=8,height=6,dpi=300,units="cm")
psave("Cross1",p3)
psave("Figure2_panelc",plots[[2]][[3]])
psave("Figure2_paneld",ppp1)
psave("Figure2_panele.png",ppp2)
psave("Figure1E",plots[[1]][[5]] + ylim(-0.2,1.2) + xlim(15,45) + theme(legend.position="none"),device="png",dpi=300,width=16,height=12)
# psave("Figure1_Aa",figures[[1]]$A,dpi=600)
# psave("Figure1_Ab",figures[[1]]$A,dpi=300)
# psave("Figure1_Ac",figures[[1]]$A,dpi=600,width=2)
# psave("Figure1_Ad",figures[[1]]$A,dpi=600,width=4)
# psave("Figure1_Ae",figures[[1]]$A,dpi=600,width=8)
# psave("Figure1_Af",figures[[1]]$A,dpi=600,width=8,height=6)
# psave("Figure1_Aa",figures[[1]]$A,dpi=300,width=16,height=12)
# psave("Figure1_Ab",figures[[1]]$A+theme_plot(),dpi=300,width=16,height=12)
# psave("Figure1_Ah",figures[[1]]$A,dpi=300,width=32,height=24)
# psave("Figure1_Ai",figures[[1]]$A,dpi=600,width=32,height=24)


```

```{r Figure_Gibbs}

colplot= FALSE
single <- TRUE
analysis <- "Gibbs"
Gibbs <- read.delim("/scicore/home/sinnreic/kinter/Rprojects.old/rPCR/documents/extdata/Gibbs energy probe.txt")

exclude <- c(5,8,9)
if (single) {
Gibbs <- Gibbs[!Gibbs$Probe.ID %in% exclude,]
Gibbs <- Gibbs[order(Gibbs$gene),]
}

Gibbs$y <- (Gibbs$strand=="anti")*0.01
Gibbs$y <- Gibbs$y + (Gibbs$isoform=="exclusion")*0.03
Gibbs$y <- (1 - Gibbs$y) 

Gibbs$x <- if (single) rep(c(1,2,4.5,5.5,8,9),each=4) else  rep(c(1,2,3.5,4.5,5.25,6.75,7.75,8.5,9.25),each=4)

dpi = 75
psize = dpi/75
figures[[analysis]] <-if (colplot) ggplot(Gibbs, aes(x=x, y=y, size=-dG,color= gene , alpha=-dG)) else ggplot(Gibbs, aes(x=x, y=y, size=-dG*psize)) # , alpha=-dG)) 
figures[[analysis]] <- figures[[analysis]] +
  geom_point() + theme_rpcr() + 
if (colplot) figures[[analysis]] <- figures[[analysis]] +    scale_colour_manual(values = rpcr_col_target[1:3])+
  # scale_fill_distiller(type = "seq",
  #                       direction = 1,
  #                       palette = "Greys",guide = "none") + 
  # scale_color_discrete(values = rpcr_col_target[1:3], guide = "none") +
  scale_y_continuous(limits = c(0, 5))  +  theme_void()+
    theme(legend.position="bottom")+ 
    scale_size_continuous(range = c(0,10))

figures[[analysis]] <- figures[[analysis]] +
theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())




figures[[analysis]]

ggsave(paste0(fpath,analysis,".png"), figures[[analysis]],width=4, dpi=300)

```

```{r}


df <- 
  tidyr::pivot_longer(import,
    cols = c("FA","FB","bFA","bFB"),   # columns to reshape
    names_to = "variable",       # new column for former column names
    values_to = "value"          # new column for values
  )


df <- df[df$sample %in% c("0", "100"), ]

df$info <- "template"
a <- "FA"
b <- "FB"
df <- df[df$variable %in% c(a,b),]
df$cat <- FALSE
df[df$sample==0,]$cat <- df[df$sample==0,]$variable==b
df[df$sample==100,]$cat <- df[df$sample==100,]$variable==a

df$cat <- c("correct","cross")[2-1*df$cat]


plot <- ggplot(data=df,aes(x=cycle,y=value,group=interaction(well,variable),col=cat)) + geom_line() + ggh4x::facet_grid2(target~sample,axes = "all")
plot <- plot + theme(axis.line=element_line(colour = "black", size = 1),
                    strip.background = element_blank(),
                     axis.title.x=element_blank(),
                     axis.title.y=element_blank(),
                    panel.background=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank()) +
  scale_colour_manual(values=c("black","red"))

plot1 <- plot

plot <- ggplot(data=df,aes(x=cycle,y=value,group=interaction(well,variable),col=variable)) + geom_line() + ggh4x::facet_grid2(target~sample,axes = "all")
plot <- plot + theme(axis.line=element_line(colour = "black", size = 1),
                    strip.background = element_blank(),
                     axis.title.x=element_blank(),
                     axis.title.y=element_blank(),
                    panel.background=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank()) +
  scale_colour_manual(values=c(rpcr_col_probe[1],rpcr_col_probe[3]))

ggsave(paste0(fpath,"Cross2.png"), plot, width=6,height=6, dpi=300)
if (exists("p3")) {
  p <- ggpubr::ggarrange(plot,p3,common.legend = TRUE, legend="bottom",widths = c(5,2))
}
ggsave(paste0(fpath,"Cross3.png"), p, width=6,height=6, dpi=300)
ggsave(paste0(fpath,"Cross2.png"), p3, width=3, height=6,dpi=300)

```

```{r}

analysis <- "PCR data collection"

# rpcr_plotCalibSummary(import)
figures[[analysis]] <- rpcr_plotCalibSummary(import,colplot=TRUE)
figures[[analysis]]





```

```{r mergely2, fig.width = 14, fig.height = 20, fig.dpi = 600}
#### rel cycle 0

analysis <- "Calib overview"

plots <- rpcr_plotCalibSummary(import,
                              type="all",
                              colplot = colplot,
                              color_target = rpcr_col_target,
                              color_sample = rpcr_col_sample)


p1 <- rpcr_plotCalib(data=import,
                    qual = "",
                    calibrate = FALSE,
                    model = "single",
                    rtype = "ip",
                    confidence = FALSE,
                    error = FALSE,
                    smooth = FALSE,
                    limit = c(NA,NA),
                    ratio = 1,
                    colplot = colplot,
                    color_target = rpcr_col_target,
                    color_sample = rpcr_col_sample)

p2 <- rpcr_plotCalib(import,
                    qual = "",
                    calibrate = FALSE,
                    model = "single",
                    rtype = "ip",
                    confidence = FALSE,
                    error = FALSE,
                    smooth = FALSE,
                    limit = c(NA,NA),
                    ratio = 1,
                    signal = "bF",
                    colplot = colplot,
                    color_target = rpcr_col_target,
                    color_sample = rpcr_col_sample)                    

part1 <- ggpubr::ggarrange(plots[[1]][[1]],
                   plots[[2]][[1]],
                   plots[[3]],
                   plots[[4]],
                   ncol = 2,
                   nrow = 2,
                   labels = c("A","B","C","D"),
                   legend ="bottom",
                   common.legend=TRUE) # ,hjust = 2)

 part2 <- ggpubr::ggarrange(p1[[3]],
                    p2[[3]],
                    ncol = 2,
                    labels = c("E","F"),
                    legend ="bottom",
                    common.legend=TRUE) # ,hjust = 5)
 
figures[[analysis]]$A <-  ggpubr::ggarrange(part1,
                      part2,
                      ncol = 1,
                      nrow = 2,
                      heights = c(2,1),
                      legend ="bottom",
                      common.legend = TRUE)
 
 example1 <- rpcr_fit(import[import$well == "E7", c("FA", "FB")],
                    model = "single",
                    rtype = "ip",
                    plot = TRUE,
                    aslist = TRUE,
                    line = TRUE)$plot
 
 example2 <- rpcr_fit(import[import$well == "E7", c("bFA", "bFB")],
                    model = "single",
                    rtype = "ip",
                    plot = TRUE,
                    aslist = TRUE,
                    line = TRUE)$plot

 part3 <- ggpubr::ggarrange(example1[[1]],example1[[2]],example2[[1]],example1[[2]],
                      ncol = 4,
                      nrow = 1,
                      legend ="bottom",
                      common.legend = TRUE)
 
 
figures[[analysis]]$B <-  ggpubr::ggarrange(part3,
                      part1,
                      ncol = 1,
                      nrow = 2,
                      heights = c(2,5),
                      legend ="bottom",
                      common.legend = TRUE)
 
 
 
 
figures[[analysis]]$A
figures[[analysis]]$B
 
```

```{r}
analysis <- "Calibration I"
# import <- import[order(import$target),]
settings <- list(qual="",
                 model = "single",
                 rtype = "ip",
                 confidence = FALSE,
                 error = FALSE,
                 smooth = FALSE,
                 limit = c(-1,1),
                 ratio = 1,
                 colplot = colplot,
                 color_target = rpcr_col_target,
                 color_sample = rpcr_col_sample)

# df0 <- do.call(rpcr_calibrate,c(list(data=import,details=TRUE),settings))$summary

p1 <- do.call(rpcr_plotCalib,c(list(import,calib = TRUE),settings)) 
p2 <- do.call(rpcr_plotCalib,c(list(import, var="cPSI"),settings)) 
p3 <- do.call(rpcr_plotCalib,c(list(import, calib = TRUE, signal = "deltaRn"),settings))
p4 <- do.call(rpcr_plotCalib,c(list(import, var = "cPSI", signal = "deltaRn"),settings)) 

figures[[analysis]] <-  ggpubr::ggarrange(p1[[3]],p2[[3]],p3[[3]],p4[[3]],
                      ncol = 4,
                      nrow = 1,
                      legend ="bottom",
                      common.legend = TRUE)
figures[[analysis]]

```

```{r}

targets <- unique(import$target)
datatypes <- c("Rn","deltaRn")
derivatives <- c("0D","1D")
crange <- c(-10:15)
calibrations <-  c("USI","cPSI")

details <- setNames(lapply(datatypes, function(datatype)
             setNames(lapply(derivatives, function(derivative)
               setNames(lapply(crange,function(cycle) 
                 rpcr_calibrate(import,
                               qual = "",
                               output = "details",
                               signal = datatype,
                               derivative = if(derivative=="1D") TRUE else FALSE,
                               rtype = "ip", # method to define range
                               roff = c(cycle,cycle), # offset from range defined by rtype
                               model = "single" # fitting method: either linear or 2nd degree polynomial (poly)
                   )
                 ), paste("c ",crange))
             ), derivatives)
               ), datatypes)


results_data <-   do.call(rbind.data.frame, lapply(datatypes, function(datatype)
                    do.call(rbind.data.frame, lapply(derivatives, function(derivative)
                      do.call(rbind.data.frame, lapply(crange,function(cycle) 
                        cbind(datatype = datatype,
                              derivative = derivative,
                              cycle = cycle,
                              details[[datatype]][[derivative]][[paste("c ",cycle)]]$summary
                        )
                        
                    ))
                    ))
                    ))




results_qc <- do.call(rbind.data.frame, lapply(datatypes, function(datatype)
                do.call(rbind.data.frame, lapply(derivatives, function(derivative)
                  do.call(rbind.data.frame, lapply(crange,function(cycle) 
                    cbind(datatype = datatype,
                       cycle = cycle,
                       derivative = derivative,
                       details[[datatype]][[derivative]][[paste("c ",cycle)]]$calib$summary$qc,
                       details[[datatype]][[derivative]][[paste("c ",cycle)]]$calib$summary$calib[,-1]
                    )
               ))
             ))
))
```

```{r}


df <- melt(results_data,id.vars=c("datatype","derivative","cycle","well","sample","target","tPSI"), measure.vars=c("USI","cPSI","delta"))
df <- df[!(df$datatype=="deltaRn" & df$derivative == "1D"),]
df <- df[order(df$datatype,decreasing=TRUE),]


ppp1 <- ggplot() + theme_rpcr() + geom_hline(yintercept = unique(df$tPSI), colour="grey50", linetype = "longdash") + 
                     geom_jitter(data = df[!(df$variable=="delta"),], aes(x = cycle, y = value, fill = tPSI), shape=21, col = "grey50",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     # paste(calibration,"using",datatype), title = target)  +
                      scale_y_continuous(breaks=unique(df$tPSI),limits=c(-0.1,1.1)) +
       ggh4x::facet_grid2( target ~ derivative + factor(datatype,c("Rn","deltaRn")) + variable ,
                            axes = "x") 
# theme(strip.background = element_blank(), strip.text = element_blank())
                           # strip = ggh4x::strip_themed(background = element_blank())


ppp1b <-ggplot() + theme_rpcr() + geom_hline(yintercept = unique(df$tPSI), colour="grey50", linetype = "longdash") + 
                     geom_jitter(data = df[!(df$variable=="delta"),], aes(x = cycle, y = value, fill = tPSI), shape=21, col = "grey50",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     # paste(calibration,"using",datatype), title = target)  +
                      scale_y_continuous(breaks=unique(df$tPSI),limits=c(-0.1,1.1)) +
       facet_grid( target ~ derivative + factor(datatype,c("Rn","deltaRn")) + variable)  + theme(strip.text.x = element_blank(),strip.text.y = element_blank(),strip.background = element_blank())

ppp2 <- ggplot() + theme_rpcr() +
                     geom_jitter(data = df[df$variable=="delta",], aes(x = cycle, y = value, fill = tPSI), shape=21, col = "grey50",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     # paste(calibration,"using",datatype), title = target)  +
                      scale_y_continuous(limits=c(-0.1,0.1)) +
       ggh4x::facet_grid2(derivative + factor(datatype,c("Rn","deltaRn"))  ~ target + variable,
                            switch = "y",
                            axes = "x") 
# theme(strip.background = element_blank(), strip.text = element_blank())
                           # strip = ggh4x::strip_themed(background = element_blank())




df <- results_data

plots <- setNames(lapply(datatypes,function(datatype)
              setNames(lapply(derivatives,function(derivative)
                 setNames(lapply(targets, function(target)
                 setNames(lapply(calibrations,function(calibration) {
                  ggplot() + theme_rpcr() + geom_hline(yintercept = unique(df$tPSI), colour="grey30", linetype = "longdash") + 
                     geom_jitter(data = df[df$datatype == datatype & df$target== target & df$ derivative== derivative , ], aes(x = cycle, y = get(calibration), fill = tPSI), shape=21, col = "grey30",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     labs(y = "",title= if(calibration=="USI") "uncalibrated" else "calibrated") +
                     # paste(calibration,"using",datatype), title = target)  +
                      scale_y_continuous(breaks=unique(df$tPSI),limits=c(-0.1,1.1))
                 }), calibrations)
                 ), targets)
               ), datatypes)
 ), derivatives)


plots1 <- setNames(lapply(datatypes,function(datatype)
              setNames(lapply(derivatives,function(derivative)
                 setNames(lapply(targets, function(target) {
                  ggplot() + theme_rpcr() + 
                     geom_jitter(data = df[df$datatype == datatype & df$target== target & df$ derivative== derivative , ], aes(x = cycle, y = delta, fill = tPSI), shape=21, col = "grey30",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     labs(y = "") +
                     # paste(calibration,"using",datatype), title = target)  +
                       scale_y_continuous(limits=c(-1,1))
                 }), targets)
               ), datatypes)
 ), derivatives)

plots2 <- lapply(derivatives,function(derivative)
               lapply(datatypes,function(datatype)
                 lapply(targets, function(target)
                 lapply(calibrations,function(calibration) {
                  ggplot() + theme_rpcr() + geom_hline(yintercept = unique(df$tPSI), colour="grey30", linetype = "longdash") + 
                     geom_jitter(data = df[df$datatype == datatype & df$target== target & df$ derivative== derivative , ], aes(x = cycle, y = get(calibration), fill = tPSI), shape=21, col = "grey30",alpha=0.7, size = 1,width=0.2,height=0) +
                     scale_fill_gradientn(colours = rpcr_col_probe) +
                     labs(y = "",title= if(calibration=="USI") "uncalibrated" else "calibrated") +
                     # paste(calibration,"using",datatype), title = target)  +
                      scale_y_continuous(breaks=unique(df$tPSI),limits=c(-0.1,1.1))
                 })
                 )
               )
 )


   
dfx <- melt(results_qc, id.vars=c("datatype","cycle","target","derivative"),
           measure.vars=c("RSSfit","RSEfit","RSScpsi","RSEcpsi","MSDfit","MSDcpsi","facRSE","facMSD"))
selection <- c("MSDfit","MSDcpsi")
df <- dfx[!(dfx$datatype=="deltaRn"&dfx$derivative=="1D")&dfx$variable %in% selection,]
parameters <- unique(df$variable)
plots1MSE <- setNames(lapply(parameters,function(parameter)
                 setNames(lapply(targets, function(target) {
                  ggplot(data = df[df$target== target&df$variable==parameter , ]) + theme_rpcr() + 
                     geom_line(aes(x = cycle, y = value, col = interaction(datatype,derivative)), size = 1) +
                         coord_cartesian(ylim=c(0,5)) +labs(y=paste(target,parameter))
                   # +
                   #   scale_fill_gradientn(colours = rpcr_col_probe) +
                   #   labs(y) 
                     # paste(calibration,"using",datatype), title = target)  +
                    #   scale_y_continuous(breaks=unique(df$tPSI),limits=c(-1,1))
                 }), targets)
               ), parameters)
selection <- c("RSEfit","RSEcpsi")
df <- dfx[!(dfx$datatype=="deltaRn"&dfx$derivative=="1D")&dfx$variable %in% selection,]
parameters <- unique(df$variable)
plots1RSE <- setNames(lapply(parameters,function(parameter)
                 setNames(lapply(targets, function(target) {
                  ggplot(data = df[df$target== target&df$variable==parameter , ]) + theme_rpcr() + 
                     geom_line(aes(x = cycle, y = value, col = interaction(datatype,derivative)), size = 1) +
                       coord_cartesian(ylim=c(0,0.05)) + labs(y=paste(target,parameter))
                   # +
                   #   scale_fill_gradientn(colours = rpcr_col_probe) +
                   #   labs(y) 
                     # paste(calibration,"using",datatype), title = target)  +
                    #   scale_y_continuous(breaks=unique(df$tPSI),limits=c(-1,1))
                 }), targets)
               ), parameters)

plotMix1 <- ggpubr::ggarrange(plotlist=unlist(plots1MSE, recursive = FALSE),ncol=3,nrow=2)
plotMix2 <- ggpubr::ggarrange(plotlist=unlist(plots1RSE, recursive = FALSE),ncol=3,nrow=2)
plotMix3 <- ggpubr::ggarrange(plotlist=c(unlist(plots1MSE, recursive = FALSE),unlist(plots1RSE, recursive = FALSE)),ncol=3,nrow=4)
plotMix4 <- ggpubr::ggarrange(plotlist=unlist(plots1RSE, recursive = FALSE)[4:6],ncol=3,nrow=1)





finish12 <- ggpubr::ggarrange(plots$`0D`$Rn$KIF13A$USI,plots$`0D`$Rn$KIF13A$cPSI,plots$`0D`$Rn$NUMA1$USI,plots$`0D`$Rn$NUMA1$cPSI,plots$`0D`$Rn$SYNE1$USI,plots$`0D`$Rn$SYNE1$cPSI,
                      plots$`0D`$deltaRn$KIF13A$USI,plots$`0D`$deltaRn$KIF13A$cPSI,plots$`0D`$deltaRn$NUMA1$USI,plots$`0D`$deltaRn$NUMA1$cPSI,plots$`0D`$deltaRn$SYNE1$USI,plots$`0D`$deltaRn$SYNE1$cPSI,
                      plots$`1D`$Rn$KIF13A$USI,plots$`1D`$Rn$KIF13A$cPSI,plots$`1D`$Rn$NUMA1$USI,plots$`1D`$Rn$NUMA1$cPSI,plots$`1D`$Rn$SYNE1$USI,plots$`1D`$Rn$SYNE1$cPSI,
                      ncol=6,
                      nrow=3,
                      legend ="bottom",
                      common.legend = TRUE)

finish13 <- ggpubr::ggarrange(plotlist <- do.call(c, unlist(plots2, recursive=FALSE)),
                      ncol=6,
                      nrow=3,
                      legend ="bottom",
                      common.legend = TRUE)





ppp1
ppp2

plotMix3

finish12


```

```{r Figure_1a, fig.width=8,fig.height=12}
yzoom <- c(-0.05,0.05)
xzoom <- c(-5,10)

qc <- "RSEcpsi"
results <- results_data
res_qc1 <- results_qc

results <- results[!(results$derivative=="1D"&results$datatype=="deltaRn"),]
results_qc <- results_qc[!(results_qc$derivative=="1D"&results_qc$datatype=="deltaRn"),]


failed <- results_qc[is.na(results_qc[[qc]]),]
failed$failed <- as.factor(interaction(failed$datatype,failed$target,failed$derivative))

results <- results_data

nullplot <- ggplot() + theme_void()

p1a <- rpcr_plotCalibCycle(results[results$derivative=="0D"&results$datatype=="Rn",],title="Rn") + theme_rpcr() + theme(plot.margin = unit(c(0.1,0.2,0.1,0.2), 'lines'))
p1b <- p1a + coord_cartesian(ylim=yzoom,xlim=xzoom)+ ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines'))

p2a <- rpcr_plotCalibCycle(results[results$derivative=="0D"&results$datatype=="deltaRn",],title="delta Rn") + theme_rpcr() + theme(plot.margin = unit(c(0.1,0.2,0.1,0.2), 'lines'))
p2b <- p2a + coord_cartesian(ylim=yzoom,xlim=xzoom)+ ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines'))

p3a <- rpcr_plotCalibCycle(results[results$derivative=="1D"&results$datatype=="Rn",],title="Rn") + theme_rpcr() + theme(plot.margin = unit(c(0.1,0.2,0.1,0.2), 'lines'))
p3b <- p3a + coord_cartesian(ylim=yzoom,xlim=xzoom)+ ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines'))

p5a <-   ggplot(data=na.omit(results_qc)) +
  geom_line(aes(x=cycle,y=get(qc), col = target, linetype = interaction(datatype,derivative)),linewidth=0.5)+
    geom_point(aes(x=cycle,y=get(qc), col = target, shape = interaction(datatype,derivative)),size=2) + #,group=interaction(gene,datatype))) +
    scale_shape_manual(values=c(20,21,22)) +
    labs( y = "average delta")  + theme_rpcr() + scale_y_log10() + scale_colour_brewer(palette = "Dark2")



p5b <- ggplot(failed) + geom_point(aes(x=cycle,y=failed, col = target, group = interaction(target,datatype),shape = interaction(datatype,derivative))) + # ,width=0,height=0.1) +
        scale_shape_manual(values=c(20,21,22,23)) +
        theme_rpcr() +
        theme(
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(),
        axis.line.x=element_blank(),
        axis.title.x=element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank()
        ) + scale_colour_brewer(palette = "Dark2") + xlim(min(crange),max(crange))




p4a  <- ggpubr::ggarrange(p1a, p1b,p2a, p2b,p3a,p3b, ncol=2, nrow=3,  legend= "none", labels=c("A","B","C","D"),align="v")
p4b  <- ggpubr::ggarrange(p5b + theme(plot.margin = unit(c(0,0,0,0), 'lines')),p5a + theme(plot.margin = unit(c(0,0,0.2,0), 'lines')), ncol=1, heights = c(1,5), common.legend = TRUE, legend="bottom",align="v")

p_legend_a <- as_ggplot(get_legend(p1a
                                   + guides(colour = guide_legend(nrow = 1))
                                   + labs(color = "tPSI")
                                   ))
p_legend_b1 <- as_ggplot(get_legend(p3b + guides(shape = FALSE) + theme(plot.margin = unit(c(0,0,0,0), 'lines')) ))
p_legend_b2 <- as_ggplot(get_legend(p3b + guides(color = FALSE) + theme(plot.margin = unit(c(0,0,0,0), 'lines')) ))

p4 <- ggpubr::ggarrange(nullplot,p4a,p_legend_a,
                # ggpubr::ggarrange(nullplot, p_legend_a, nullplot, nrow = 1, widths = c(0.5,1,0.5), legend = "none"),
                nullplot,
                ggpubr::ggarrange(nullplot, nullplot, nullplot, nrow = 1, widths = c(0.5,1,0.5), legend = "none",labels=c("","E","")),
                ggpubr::ggarrange(nullplot, p4b, nullplot, nrow = 1, widths = c(0.5, 1, 0.5), legend = "right"),
                nullplot,
                p_legend_b1,
                p_legend_b2,
                nrow=9, heights = c(0.1,3,0.1,0.05,0.15,1,0.05,0.1,0.1)) # ,legend="none"

Figure_1 <- annotate_figure(p4, top = text_grob("Calculation with single cycle relative to inflection point",
               color = "black", face = "bold", size = 14))




Figure_1
```

```{r}

FSYNE1 <- data.frame(import[import$target=="SYNE1",]  %>% group_by(rcycle) %>% summarise(total=mean(bFtotal)))
Fall <- data.frame(import  %>% group_by(rcycle,target) %>% summarise(total=mean(bFtotal)))

names(FSYNE1)[1] <- "cycle"
names(Fall)[1] <- "cycle"

# melt(results_qc[results_qc$target=="SYNE1",],
#            value_name="value",
#            measure.vars =c("a","ri","re","g"),
#            id = c("cycle","target","sample","well"))

ttt1 <- merge(results_qc[results_qc$target=="SYNE1",],FSYNE1,by="cycle")
ttt1 <- merge(results_qc,Fall,by=c("cycle","target"))

ggplot(data=ttt1) + geom_point(aes(x=total,y=a,col=target,group=interaction(target,derivative,datatype)))
ggplot(data=ttt1[ttt1$datatype=="Rn"&ttt1$derivative=="1D",]) + geom_point(aes(x=total,y=a,col=target,group=interaction(target,derivative,datatype)))

datatypes <- "deltaRn"

ttt2 <- ttt1[ttt1$datatype==datatypes&ttt1$derivative=="1D",]
ttt2 <- ttt1
# ttt2$total <- log(ttt2$total)
sel <- ttt2[ttt2$cycle %in% c(-2:1),]

ggplot(data=ttt2) + geom_line(aes(x=total,y=a,col=interaction(target,derivative,datatype)))+ geom_point(aes(x=total,y=a,col=interaction(target,derivative,datatype))) + geom_point(data=sel,aes(x=total,y=a,group=interaction(target,derivative,datatype)),col=1) + ylim(0,5)




```

```{r}


df <- results_data
  

df$signal <- interaction(df$datatype,df$derivative)
df <- df[df$signal!="deltaRn.1D",]


df$signal <- factor(df$signal, levels=c("Rn.0D","deltaRn.0D","Rn.1D"), labels=c("Rn","deltaRn","derRn")) 
  jwidth <- unique(df[,"cycle"])
  jwidth <- jwidth[order(jwidth)]
  jwidth <- mean(jwidth[2:(length(jwidth))]- jwidth[1:(length(jwidth)-1)])/5


    plot <- ggplot(data=df) +
      # geom_rect(data = data.frame(target= unique(df[order(df$target),"target"]),x=1), aes(fill = target), col = NA, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, alpha = 0.1) +

        theme_rpcr() +  scale_colour_gradientn(colours = rpcr_col_probe)  


  plot1 <- plot + 
    ggh4x::facet_grid2(target ~ signal, switch = "y", strip = ggh4x::strip_themed(background_y = ggh4x::elem_list_rect(fill = rpcr_col_target)),axes = "x") +
    geom_boxplot(aes(x=cycle,y=delta,group=cycle),outlier.shape = NA) +
    geom_jitter(aes(x=cycle,y=delta, group=interaction(cycle,tPSI),col=tPSI), # col= if (colplot) as.factor(tPSI) else "black",
                             width=jwidth,
                             alpha=0.5,
                             size=1) 
  
  
  plot1b <- plot + 
    facet_grid(target ~ signal) +
        geom_boxplot(aes(x=cycle,y=delta,group=cycle),outlier.shape = NA) +
    geom_jitter(aes(x=cycle,y=delta, group=interaction(cycle,tPSI),col=tPSI), # col= if (colplot) as.factor(tPSI) else "black",
                             width=jwidth,
                             alpha=0.5,
                             size=1) +
   theme(strip.text.x = element_blank(),strip.text.y = element_blank(),strip.background = element_blank())

  
  plot2 <- ggplot(data=df) +
    theme_rpcr() +  scale_colour_gradientn(colours = rpcr_col_probe)  +
    ggh4x::facet_grid2(target ~ signal,  axes = "all")+
    geom_violin(aes(x=cycle,y=delta,group=cycle)) + #,fill="grey90") +
    geom_jitter(aes(x=cycle,y=delta, group=interaction(cycle,tPSI),col=tPSI), # col= if (colplot) as.factor(tPSI) else "black",
                             width=jwidth,
                             alpha=0.5,
                             size=1)+
    coord_cartesian(ylim=yzoom,xlim=c(-1.1,7.1)) +
    ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines')) + scale_x_continuous(breaks=-1:7)




```

```{r}


df <- results_qc
  


df$signal <- interaction(df$datatype,df$derivative)
df <- df[df$signal!="deltaRn.1D",]
df$signal <- factor(df$signal, levels=c("Rn.0D","deltaRn.0D","Rn.1D"), labels=c("Rn","deltaRn","derRn")) 


failed <- df[is.na(df[["RSEcpsi"]]),]
failed$failed <- as.factor(interaction(failed$datatype,failed$target,failed$derivative))

qc <- "RSEcpsi"


p5a <-   ggplot(data=na.omit(df)) +
  geom_line(aes(x=cycle,y=get(qc), col = target, linetype = interaction(datatype,derivative)),linewidth=0.5)+
    geom_point(aes(x=cycle,y=get(qc), col = target, shape = interaction(datatype,derivative)),size=2) + #,group=interaction(gene,datatype))) +
    scale_shape_manual(values=c(20,21,22)) + facet_grid(target~.)+
    labs( y = "average delta")  + theme_rpcr() + scale_y_log10() + scale_colour_brewer(palette = "Dark2")



p5b <- ggplot(failed) + geom_point(aes(x=cycle,y=failed, col = target, group = interaction(target,datatype),shape = interaction(datatype,derivative))) + # ,width=0,height=0.1) +
        scale_shape_manual(values=c(20,21,22,23)) +
        theme_rpcr() +
        theme(
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(),
        axis.line.x=element_blank(),
        axis.title.x=element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank()
        ) + scale_colour_brewer(palette = "Dark2") + xlim(min(crange),max(crange))
```

analyzing treshold range

```{r range1, echo=FALSE}

analysis <- "tres single range"

df <- import

datalist <- "Rn"
range1_list <- 1
range2_list <- seq(0,0.99,0.01)




res[[analysis]]$details <-   setNames(lapply(datalist, function(datatype)
               setNames(lapply(range1_list,function(range1)
               setNames(lapply(range2_list,function(range2) {
                 print(paste(datatype,range1,range2))
                 rpcr_calibrate(df,
                               qual="",
                               output = "details",
                               signal = datatype,
                               tres = range2,     # treshold value to define dynamic range
                               rtype = "dynamic", # method to define range
                               roff = 0, # offset from range defined by rtype
                               model = "linear" # fitting method: either linear or 2nd degree polynomial (poly)
                   ) 
                 }), paste0("R_",range2_list))
               ), paste0("R_",range1_list))
               ), datalist)


res[[analysis]]$results_data <-   do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)               
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                              res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                              )
                        ))
                      ))
                    ))

res[[analysis]]$results_qc <-   do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)               
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                               res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                       )
                        ))
                      ))
                    ))
             



```

```{r}
analysis <- "tres single range"

df <- res[[analysis]]$results_qc
results_qc_na <- results_qc %>% replace(is.na(.), 1)

var <- "RSEcpsi"

top <- do.call(rbind.data.frame, lapply(genelist, function(target)
             do.call(rbind.data.frame, lapply(datalist, function(datatype){
               temp <- df[df$target==target&df$datatype==datatype, ]
               temp[which.min(temp$RSEcpsi),]
             }
             ))
           ))

top <- merge(dcast(df, datatype + range1 + range2 ~ target,value.var=var),dcast(df, datatype + range1 + range2 ~ .,value.var=var,fun=mean))
names(top)[names(top)=="."] <- "mean"
top$mean <- top$mean*100

print(top)
print(top[which.min(top$mean),])

figures[[analysis]] <- ggplot(df, aes(x=range2,y=get(var),col=target)) + 
  geom_line()
figures[[analysis]]

```

analyzing treshold range both sides

```{r range10, echo=FALSE}

analysis <- "tres double range"

df <- import

datalist <- "Rn"
range1_list <- seq(0,0.99,0.05)
range2_list <- seq(0,0.99,0.05)



res[[analysis]]$details <-   setNames(lapply(datalist, function(datatype)
               setNames(lapply(range1_list,function(range1){               
               print(paste(datatype,range1))
               setNames(lapply(range2_list,function(range2) {
                 rpcr_calibrate(df,
                               qual="",
                               output = "details",
                               signal = datatype,
                               tres = c(range1,range2),     # treshold value to define dynamic range
                               rtype = "dynamic", # method to define range
                               roff = 0, # offset from range defined by rtype
                               model = "linear" # fitting method: either linear or 2nd degree polynomial (poly)
                   ) 
                 }), paste0("R_",range2_list))}
               ), paste0("R_",range1_list))
               ), datalist)


res[[analysis]]$results_data <- do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)                
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                              res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                              )
                        ))
                      ))
                    ))

res[[analysis]]$results_qc <-   do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)               
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                               res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                       )
                        ))
                      ))
                    ))
             
res[[analysis]]$details <- NULL


```

figures\[\[analysis\]\]

```{r}

analysis <- "tres double range"

df <- res[[analysis]]$results_qc
results_qc_na <- results_qc %>% replace(is.na(.), 1)

var <- "RSEcpsi"

top <- do.call(rbind.data.frame, lapply(genelist, function(target)
             do.call(rbind.data.frame, lapply(datalist, function(datatype){
               temp <- df[df$target==target&df$datatype==datatype, ]
               temp[which.min(temp$RSEcpsi),]
             }
             ))
           ))

top <- merge(dcast(df, datatype + range1 + range2 ~ target,value.var=var),dcast(df, datatype + range1 + range2 ~ .,value.var=var,fun=mean))
names(top)[names(top)=="."] <- "mean"
top$mean <- top$mean*100

print(top)
print(top[which.min(top$mean),])


figures[[analysis]] <- lapply(genelist, function(target){
scaly <- quantile(df[df$target==target,var],0.1,na.rm=TRUE)
  ggplot(df[df$target==target,], aes(range1, range2, fill= get(var))) + 
  geom_tile() +
  scale_fill_viridis() 
  # scale_fill_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
  #   ifelse(x<scaly, 
  #          scales::rescale(x,
  #                          to = to,
  #                          from = c(min(x, na.rm = TRUE), scaly)),
  #          1)}) 
  
})
figures[[analysis]]

```

Analyzing IP range

```{r range11, echo=FALSE}

analysis <- "IP range"

df <- import

datalist <- "Rn"
range1_list <- seq(-20,20,1)
range2_list <- seq(-20,20,1)




res[[analysis]]$details <-  setNames(lapply(range1_list,function(range1){               
               print(range1)
               setNames(lapply(range2_list,function(range2) {
                 rpcr_calibrate(df,
                               qual="",
                               output = "details",
                               rtype = "ip", # method to define range
                               roff = c(range1,range2), # offset from range defined by rtype
                               model = "linear" # fitting method: either linear or 2nd degree polynomial (poly)
                   )
                 }), paste0("R_",range2_list))}
               ), paste0("R_",range1_list))
               


res[[analysis]]$results_data <- do.call(rbind.data.frame, lapply(range1_list, function(range1)                
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(range1 = range1,
                              range2 = range2,
                              res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                              )
                        ))
                      ))
                    

res[[analysis]]$results_qc <-   
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)               
                      do.call(rbind.data.frame, lapply(range2_list, function(range2) 
                        cbind(range1 = range1,
                              range2 = range2,
                               res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                       )
                        ))
                      ))


paraQC <- "RSEcpsi"
res[[analysis]]$results_qc2 <- 
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)               
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)  {
                    settings <- standardSettings
                    settings[,c("rtype","roff")] <- c("ip:ip",paste(c(range1,range2),collapse=":"))
                    temp <-res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                    rownames(temp) <- temp$target
                    temp2 <- data.frame(t(temp))[paraQC,]
                    cbind(settings,mean = mean(temp[,paraQC]),temp2)  
                    }
               ))
             ))


                    
             
res[[analysis]]$details <- NULL


```

```{r, fig.width = 3, fig.height = 9, fig.dpi = 600}

analysis <- "IP range"

df <- res[[analysis]]$results_qc


var <- "RSEcpsi"

top <- do.call(rbind.data.frame, lapply(genelist, function(target)
             do.call(rbind.data.frame, lapply(datalist, function(datatype){
               temp <- df[df$target==target&df$datatype==datatype, ]
               temp[which.min(temp$RSEcpsi),]
             }
             ))
           ))

top <- merge(dcast(df,  range1 + range2 ~ target,value.var=var),dcast(df, range1 + range2 ~ .,value.var=var,fun=mean))
names(top)[names(top)=="."] <- "mean"
top$mean <- top$mean*100

print(top)
print(top[which.min(top$mean),])
scale_max <- max(df[[var]],na.rm=TRUE) 
scale_breaks <- c(0,0.01,0.011,0.012,0.013,0.014,0.015,0.018,0.019,scale_max)
scale_col <- rev(viridis::plasma(n = length(scale_breaks)))



df <- df[df$range2>df$range1,]

df2 <- df[!is.na(df[[var]]),]
r1 <- c(min(df2$range1,df2$range2),max(df2$range1,df2$range2))

figures[[analysis]] <- ggplot(df, aes(range1, range2, fill= cut(get(var), breaks = scale_breaks))) + 
  geom_tile(col="white") +
  scale_fill_manual(values = scale_col,na.value="grey90") + theme_rpcr() +
   facet_grid(target~.) + coord_fixed() + ylim(r1[1],r1[2])+ xlim(r1[1],r1[2])

figures[[analysis]]

ggsave(paste0(fpath,analysis,".png"), figures[[analysis]],width = 3, dpi=300, units = "in")
```

Fixed cycle IP range

```{r range_005, echo=FALSE}

analysis <- "Fixed cycle range"
df <- import

datatypes <- c("Rn","deltaRn")
derivatives <- c(FALSE,TRUE)
range1_list <- seq(-20,20,1)
range2_list <- seq(-20,20,1)

start.time <- Sys.time()
res[[analysis]]$details  <- setNames(lapply(derivatives, function(derivative)
             setNames(lapply(datatypes, function(datatype)
               setNames(lapply(range1_list,function(range1){
               print(paste(derivative,datatype,range1))
               setNames(lapply(range2_list,function(range2) {
                 rpcr_calibrate(df,
                               qual = "",
                               output = "details",
                               derivative = derivative,
                               signal = datatype,
                               tres = 0,     # treshold value to define dynamic range
                               rtype = "ip", # method to define range
                               roff = c(range1,range2), # offset from range defined by rtype
                               method = c("incl","total"),
                               model = "single") 
                 }), paste0("R_",range2_list))}
               ), paste0("R_",range1_list))
               ), datatypes)
             ),c("0D","1D"))
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

res[[analysis]]$results_data <- do.call(rbind.data.frame, lapply(derivatives, function(derivative)
             do.call(rbind.data.frame, lapply(datatypes, function(datatype)
                do.call(rbind.data.frame, lapply(range1_list,function(range1)               
               do.call(rbind.data.frame, lapply(range2_list,function(range2) {
                 cbind(derivative = derivative,
                       model = "single",
                       datatype = datatype,
                       range1 = range1,
                       range2 = range2,
                       res[[analysis]]$details[[if (derivative) "1D" else "0D"]][[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                 )}
               ))
             ))
           ))
))


res[[analysis]]$results_qc <- do.call(rbind.data.frame, lapply(derivatives, function(derivative)
                do.call(rbind.data.frame, lapply(datatypes, function(datatype)
                  do.call(rbind.data.frame, lapply(range1_list,function(range1) 
                    do.call(rbind.data.frame, lapply(range2_list,function(range2)
                      cbind(
                        datatype = datatype,
                        derivative = derivative,
                        range1 = range1,
                        range2 = range2,
                        res[[analysis]]$details[[if (derivative) "1D" else "0D"]][[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                      )
                 
               ))
             ))
           ))
))



res[[analysis]]$details <- NULL

                   
```

```{r Figure_005, fig.width=8,fig.height=12}


analysis <- "Fixed cycle range"

df <- na.omit(res[[analysis]]$results_qc)


var <- "RSEcpsi"

top <- do.call(rbind.data.frame, lapply(genelist, function(target)
        do.call(rbind.data.frame, lapply(derivatives, function(derivative)
             do.call(rbind.data.frame, lapply(datatypes, function(datatype){
               temp <- df[df$target==target&df$datatype==datatype&df$derivative==derivative, ]
               temp[which.min(temp$RSEcpsi),]
             }
             ))
           ))
))

top <- merge(dcast(df,  range1 + range2 ~ target,value.var=var),dcast(df, range1 + range2 ~ .,value.var=var,fun=mean))
names(top)[names(top)=="."] <- "mean"
top$mean <- top$mean*100

print(top)
print(top[which.min(top$mean),])


figures[[analysis]] <- lapply(genelist, function(target){
scaly <- quantile(df[df$target==target,var],0.1,na.rm=TRUE)
print(scaly)
scaly <- 0.02

print(scaly)
  ggplot(df[df$target==target,], aes(range1, range2, fill= get(var))) + 
  geom_tile() +
   scale_fill_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
   ifelse(x<scaly, 
   scales::rescale(x,
                             to = to,
                              from = c(min(x, na.rm = TRUE), scaly)),
             1)}) 
  
})












```

```{r}

analysis <- "different scenarios"
get_qc <- function(df,
                   qual="",
            signal="Rn",
            tres = 0,     # treshold value to define dynamic range
            rtype = c("full"), # method to define range
            roff = c(0,0), # offset from range defined by rtype
            method = c("incl","total"),
            model = "linear",
                   ...){
  
temp0 <- data.frame(signal=signal,rtype=paste(rtype,collapse=":"),roff=paste(roff,collapse=":"),tres=paste(tres,collapse=":"),method=paste(method,collapse=":"),model= model)
temp1 <- rpcr_calibrate(df,tres=tres,rtype=rtype,roff=roff,method=method,model=model,qual=qual,output="qc",...)
rownames(temp1) <- temp1$target

temp2 <- data.frame(t(temp1))["RSEcpsi",]

cbind(temp0,mean = mean(temp1$RSEcpsi),temp2)
}



res[[analysis]] <- rbind(
  get_qc(import,rtype=c("full","full")), 
  get_qc(import,rtype=c("full","full")),
get_qc(import,rtype=c("ip","full")),
get_qc(import,rtype=c("full","ip")),
get_qc(import,rtype=c("dynamic","full")),
get_qc(import,rtype=c("full","dynamic")),
get_qc(import,rtype=c("dynamic","dynamic")),
get_qc(import,rtype=c("dynamic","ip")),
get_qc(import,rtype=c("ip","dynamic")),
get_qc(import,rtype=c("dynamic","full"),tres=-1),
get_qc(import,rtype=c("dynamic","full"),tres=0),
get_qc(import,rtype=c("dynamic","full"),tres=0.05),
get_qc(import,rtype=c("dynamic","full"),tres=0.1),
get_qc(import,rtype=c("dynamic","full"),tres=0.15),
get_qc(import,rtype=c("dynamic","full"),tres=0.2),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0,0.1)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0,0.3)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0,0.33)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0,0.5)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0.05,0.1)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0.05,0.3)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0.05,0.33)),
get_qc(import,rtype=c("dynamic","dynamic"),tres=c(0.05,0.5))
)

res[[analysis]] 

```

rPCR simulation with different parameter values.

```{r unknown_block6}


label_size <- 10
title_size <- 14
axis_size <- 12
scale <- 1


plot_sim <- function(sim_fun="rpcr_simpsi",para,range,log=FALSE,title=NULL,ratio=2,args = list(a=1,ri=0,re=0,g=1)) {
    symb <- c(a="$\\alpha$",ri="$\\r_i$",re="$\\r_e$",g="$\\gamma$")
  
    delta <- function(x,a) {get(sim_fun)(x, a = a$a, ri = a$ri, re = a$re, g = a$g) - x}
  
  geom_layer <- function (x,args,para) {
      args[para] <- x
      geom_function(aes(color=x), fun = get(sim_fun), args = args, alpha=.7, linewidth = 0.5)
      }
  
  geom_layer2 <- function (x,args,para) {
      args[para] <- x
      geom_function(aes(color=x), fun = delta, args= list(args), alpha=.7, linewidth = 0.5)
      }
  

  layers <- lapply(range, geom_layer, args = args, para = para)
  layers2 <- lapply(range, geom_layer2, args = args, para = para)

  if (!is.null(title)) title <- title # TeX("$Modelling~different~probe~intensities~(alpha)$")
  #// combine plot and layers
print("OK0")
  p1 <- ggplot() + xlim(0, 1) + geom_function(fun= get(sim_fun), linetype = "dotted") +
      layers +
      ggtitle(title) +
      scale_color_viridis() +
      xlab(TeX("$\\psi$")) +
      ylab(TeX("$\\psi_a$")) +
      labs(color = TeX(symb[para])) +
      theme_rpcr() +
      xlim(0, 1) + ylim(0, 1) 
      # coord_fixed(ratio = 1, xlim = c(0, 1), ylim = c(0, 1),expand = FALSE) # , clip = "on")
    
  p2 <- ggplot() + xlim(0, 1) + geom_function(fun= delta, linetype = "dotted") +
      layers2 +
      # ggtitle(title) +
      scale_color_viridis() +
      xlab(TeX("$\\psi$")) +
      ylab(TeX("$\\Delta\\psi\\ [\\psi - \\psi_a$]")) +
      labs(color = TeX(symb[para])) +
      theme_rpcr() +
      scale_y_continuous(breaks=c(-0.5,0,0.5)) +
      xlim(0, 1) + ylim(-0.5, 0.5) 
      # coord_fixed(xlim = c(0, 1), ylim = c(-0.5, 0.5),expand = FALSE, clip = "on")
  print("OK1")
  if (log) {
     sim_breaks <- c(min(range),median(range),max(range))
      p1 <- p1 + scale_color_viridis(trans = "log", breaks = sim_breaks, labels = sim_breaks)
      p2 <- p2 + scale_color_viridis(trans = "log", breaks = sim_breaks, labels = sim_breaks)
  }
    print("OK2")
  merge_plots <- list(p1 + theme(plot.margin = unit(c(1,1,1,1), 'lines')),p2 + theme(plot.title = element_blank(),plot.margin = unit(c(1,1,1,1), 'lines')))
  list( ggpubr::ggarrange(plotlist=merge_plots, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="v"),p1,p2)
}
    

range <- function(mid=1,max=4,steps=.2) c(seq(mid,max,steps),1/seq(mid,max,steps))
plot_a <- plot_sim(sim_fun="rpcr_simpsi",para="a",log=TRUE,range=range(),title= TeX("$Modelling~parameter~alpha$"))
plot_re <- plot_sim(sim_fun="rpcr_simpsi",para="re",range=seq(0,1,0.05),title= TeX("$Modelling~parameter~r_e$"))
plot_ri <- plot_sim(sim_fun="rpcr_simpsi",para="ri",range=seq(0,1,0.05),title= TeX("$Modelling~parameter~r_i$"))
plot_g <- plot_sim(sim_fun="rpcr_simpsi",para="g",range=seq(1,10,0.4),title= TeX("$Modelling~parameter~gamma$"))

plot_final1 <- ggpubr::ggarrange(plot_a[[1]],plot_g[[1]],plot_ri[[1]],plot_re[[1]],ncol=2,nrow=2)
plot_final2 <- ggpubr::ggarrange(plot_a[[2]],plot_g[[2]],plot_ri[[2]],plot_re[[2]],ncol=2,nrow=2)
plot_final3 <- ggpubr::ggarrange(plot_a[[3]],plot_g[[3]],plot_ri[[3]],plot_re[[3]],ncol=2,nrow=2)
plot_final4 <- ggpubr::ggarrange(plot_final2,plot_final3,nrow=2)
plot_final5 <- ggpubr::ggarrange(plot_final2,plot_final3,ncol=2)
plot_final6 <- ggpubr::ggarrange(plot_a[[1]],plot_g[[1]],plot_ri[[1]],plot_re[[1]],ncol=4,nrow=1)


merge_theme <-       theme(
  # axis.line=element_blank(),
  # axis.text.x=element_blank(),
  axis.text.y=element_blank(),
  #    axis.ticks=element_blank(),
  #    axis.title.x=element_blank(),
      axis.title.y=element_blank())

merge_theme2 <-       theme(
  # axis.line=element_blank(),
  axis.text.x=element_blank(),
  #    axis.ticks=element_blank(),
      axis.title.x=element_blank()
   )

margin_1_a <- unit(c(1,1,0,1), 'lines')
margin_1_b <- unit(c(1,1,0,0), 'lines')
margin_2_a <- unit(c(0,0,0,0), 'lines')
margin_2_b <- unit(c(0,0,0,0), 'lines')

merge_a <- list(
  plot_a[[2]] + theme(plot.margin = margin_1_a) + merge_theme2,
  plot_a[[3]] + theme(plot.title = element_blank(),plot.margin =margin_2_a) 
  )
merge_ri <- list(
  plot_ri[[2]] + theme(plot.margin = margin_1_b) + merge_theme + merge_theme2 ,
  plot_ri[[3]] + theme(plot.title = element_blank(),plot.margin = margin_2_b) + merge_theme 
  )

merge_re <- list(
  plot_re[[2]] + theme(plot.margin = margin_1_b) + merge_theme +merge_theme2,
  plot_re[[3]] + theme(plot.title = element_blank(),plot.margin = margin_2_b) + merge_theme
  )

merge_g <- list(
  plot_g[[2]] + theme(plot.margin = margin_1_b) + merge_theme +merge_theme2,
  plot_g[[3]] + theme(plot.title = element_blank(),plot.margin = margin_2_b) + merge_theme 
  )
ratio <- 2
merge_a_f <- ggpubr::ggarrange(plotlist=merge_a, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="v")
merge_ri_f <- ggpubr::ggarrange(plotlist=merge_ri, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="v")
merge_re_f <- ggpubr::ggarrange(plotlist=merge_re, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="v")
merge_g_f <- ggpubr::ggarrange(plotlist=merge_g, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="v")

plot_final <-  ggpubr::ggarrange(merge_a_f,merge_ri_f,merge_re_f,merge_g_f,ncol=4,nrow=1,widths=c(1.29,1,1,1))

```

```{r unknown_block7, fig.width=12, fig.asp=0.435}

plot_final
```

```{r unknown_block8}

# rpcr_plotSim generates plots to compare theoretical (tPSI) and apparent PSI (aPSI) for multiple calibration parameter sets.

plot_sim2 <- function(curves) {
ratio <- 2
sim_fun <- "rpcr_simpsi"
curves <- rbind(linetype=c(rep(0.5,length(curves)-1),2),curves)
curves <- rbind(col=2:(length(curves)+1),curves)
ttt1 <- lapply(curves, function(x) print(list(x[2])))
delta <- function(x,args=list(a=1,ri=0,re=0,g=1)) {get(sim_fun)(x, a = args[[1]], ri = args[[2]], re = args[[3]], g = args[[4]]) - x}
fun_std <- function(x,args) get(sim_fun)(x, a = args[[1]], ri = args[[2]], re = args[[3]], g = args[[4]])

geom_layer1 <- lapply(curves, function (args) { geom_function(fun = fun_std, args = list(args[-c(1:2)]), alpha=.7, linewidth = args[2], col = args[1])})
geom_layer2 <- lapply(curves, function (args) { geom_function( fun = delta, args= list(args[-c(1:2)]), alpha=.7, linewidth = args[2], col = args[1])})

  p1 <- ggplot() + xlim(0, 1) + geom_function(fun= get(sim_fun), linetype = "dotted") +
      geom_layer1 +
     scale_colour_discrete() +
      # ggtitle(title) +
      # scale_color_viridis() +
      xlab(TeX("$\\psi$")) +
      ylab(TeX("$\\psi_a$")) +
      theme_rpcr() +
      xlim(0, 1) + ylim(0, 1) 
      # coord_fixed(ratio = 1, xlim = c(0, 1), ylim = c(0, 1),expand = FALSE) # , clip = "on")
    
  p2 <- ggplot() + xlim(0, 1) + geom_function(fun= delta, linetype = "dotted") +
      geom_layer2 +
      scale_colour_discrete() +
      # ggtitle(title) +
      xlab(TeX("$\\psi$")) +
      ylab(TeX("$\\Delta\\psi\\ [\\psi - \\psi_a$]")) +
      theme_rpcr() +
      xlim(0, 1) + ylim(-0.5, 0.5) 
      # coord_fixed(xlim = c(0, 1), ylim = c(-0.5, 0.5),expand = FALSE, clip = "on")
  

    
  merge_plots <- list(p1 + theme(plot.margin = unit(c(1,1,1,1), 'lines')),p2 + theme(plot.title = element_blank(),plot.margin = unit(c(1,1,1,1), 'lines')))
  ggpubr::ggarrange(plotlist=merge_plots, ncol=1,nrow=2, common.legend =TRUE,legend="bottom",heights= c(ratio,1),align="hv")
}


curves  <- data.frame(c(a=2,ri=0,re=0,g=1),c(a=1,ri=0,re=0,g=2),c(a=1,ri=0,re=0.1,g=1),c(a=2,ri=0,re=0.1,g=2))
example_1 <- plot_sim2(curves)

curves  <- data.frame(c(a=5,ri=0,re=0.0,g=1),c(a=1,ri=0.2,re=0,g=1),c(a=1,ri=0,re=0,g=4),c(a=5,ri=0.2,re=0,g=4))
example_2 <- plot_sim2(curves)

curves  <- data.frame(c(a=1/5,ri=0,re=0.0,g=1),c(a=1,ri=0.2,re=0,g=2),c(a=1,ri=0,re=0,g=4),c(a=1/5,ri=0.2,re=0,g=4))
example_3 <- plot_sim2(curves)
example_1
example_2
example_3
```

```{r SimulateSingle, fig.width=3, fig.asp=1.38}

# rpcr_plotSim generates plots to compare theoretical (tPSI) and apparent PSI (aPSI) for a single calibration parameter set.

p1 <- rpcr_plotSim(c(a=2,ri=0.1,re=0.2,g=2))
p2 <- rpcr_plotSim(c(a=0.2,ri=0,re=0,g=4))
p3 <- rpcr_plotSim(c(a=2,ri=0,re=0,g=6))
p1
p2
p3

```

```{r okok}
rpcr_plotSim(t(data.frame(a=c(1,2,3),ri=c(0,0.2,0.3),re=c(0,0.2,0.3),g=c(1,1.1,1.2))))

```

```{r}
file1 <- paste("results",Sys.Date(),collapse="_")
file2 <- paste("figures",Sys.Date(),collapse="_")
saveRDS(res, file = file1)
saveRDS(figures, file = file2)
```
