---
title: "ratioPCR Figure S6 — Mechanistic basis of equimolar drift: heteroduplex association kinetics determine proportion-dependent amplification efficiency differences"
format:
  html:
    embed-resources: true
    code-background: true
    code-fold: true
    code-tools: true
    toc: true
    toc-depth: 3
    number-sections: true
editor: visual
execute:
  warning: false
  message: false
---

# Figure legend

> **Supplementary Figure S6: Mechanistic basis of equimolar drift: heteroduplex association kinetics determine proportion-dependent amplification efficiency differences**
>
> (a) Simulation design. Single-cycle simulations were performed across a range of total template amounts (colour: dark blue = 1×10⁻⁷ M to yellow = 4×10⁻⁷ M) and starting template proportions (colour: dark blue = 0 to yellow = 1.0), using a fixed homoduplex association rate of k₁_homo = 1×10⁶ M⁻¹s⁻¹ and three different heteroduplex association rates: k₁_hetero = 1×10⁶, 6×10⁵, and 3×10⁵ M⁻¹s⁻¹. For each condition, homoduplex formation fraction, heteroduplex formation fraction, total re-annealing fraction, amplification efficiency, and new amplicon yield were calculated as a function of starting template proportion.
>
> (b) Equal association rates (k₁_hetero = 1×10⁶ M⁻¹s⁻¹). When homo- and heteroduplex formation proceed at equal rates, the fraction of molecules forming homoduplexes is minimal and heteroduplex formation is maximal at equimolar template ratios. Total re-annealing and amplification efficiency are independent of template proportion under these conditions, meaning no proportion-dependent efficiency difference arises. Amplification efficiency decreases with increasing total template concentration due to greater overall re-annealing competition.
>
> (c) Reduced heteroduplex association rate (k₁_hetero = 3×10⁵ M⁻¹s⁻¹). When heteroduplexes form less readily than homoduplexes, the total re-annealing fraction and amplification efficiency become proportion-dependent. Efficiency is highest at equimolar ratios (where heteroduplex formation is most suppressed relative to homoduplex formation) and decreases as the proportion of either template increases. This creates a template-specific efficiency difference that depends on the starting proportion. Results are shown for total templates (top two rows) and separately for template A and template B (bottom two rows), illustrating that the efficiency asymmetry affects each template differently depending on its relative abundance.
>
> (d) Proportion-dependent efficiency differences drive equimolar drift. At equal re-annealing rates, the per-template amplification efficiencies of targets A and B are identical across all proportions (left column, flat relationship). As the heteroduplex association rate is decreased to 6×10⁵ (middle) and 3×10⁵ M⁻¹s⁻¹ (right), proportion-dependent efficiency differences emerge (top row), with the efficiency ratio deviating from 1 in a proportion-dependent manner (middle row). This leads to a systematic, proportion-dependent deviation of the measured proportion from the input proportion (bottom row), producing the characteristic S-shaped bias profile seen in Figure 2G. The sole exception is the equimolar 1:1 ratio, at which both templates experience identical rehybridisation competition and no efficiency difference arises.
>
> (e) Schematic summary of the amplification inhibition mechanism. In multi-template PCR, amplification efficiency is enhanced by heteroduplex formation relative to homoduplex formation, because heteroduplexes are not substrates for extension. The less abundant template has a proportionally higher fraction of its strands in heteroduplexes, giving it a relatively higher effective efficiency compared with the more abundant template. This creates a systematic pull toward equimolar proportions that increases in magnitude as the heteroduplex association rate decreases relative to the homoduplex rate.

## Source script and scenario directory

Unlike the full-PCR simulations of Figure 2 and Supplementary Figure S5, this figure was produced from a Quarto notebook that was later purled into `.R` files and archived inside the simulation directories. Several near-identical timestamped copies exist:

| Directory | Copies | Content |
|----------------------------|---------------------|-----------------------|
| `Sim_hetero_singlecycle` | `rP_Heteroduplex.qmd_2025-09-26_2300` … `_2025-09-27_0029` (nine copies, \~3 kB) | setup and the single-cycle batch definition only |
| `Sim_Hetero_sCycle` | `…_2025-09-27_1123`, `_1124`, `_2025-09-29_1237`, `_1238` (\~9 kB) | adds the derived efficiency quantities and exploratory fits |
| `Sim_Hetero_asymetric_sCycle` | `…_2025-09-28_2015`, `_2036` (\~7 kB) | adds the asymmetric-rehybridisation variant (`rehybB_k1 = 3×10⁶`) |


The **data** all three panels read is `documents/rPCRsim/Sim_Hetero_sCycle/results/results.rds`: the notebook sets `title <- "Hetero_sCycle"` and reads `Sim_<title>/results/results.rds`, and its nested sweep already contains all three heteroduplex rates (`hetero_k1 = c(3e5, 6e5, 1e6)`). Panel b is the `hetero_k1 == 1e6` slice of that same object, not a separate simulation.

Single-cycle rather than full-PCR simulations are used here for a substantive reason: PCR is iterative, so in a 45-cycle run the per-cycle efficiency difference compounds and cannot be read off the end point. Restricting the model to `cycle_max = 1` and sweeping the *starting* amount instead isolates the per-cycle effect at each point along the amplification curve.

# Setup

```{r}
#| label: setup

pp <- here::here()
source(file.path(pp, "documents", "scripts", "Figure_settings.R"))
source(file.path(pp, "documents", "scripts", "omPCR_functions.R"))

library(ggplot2)
library(ratioPCR)
# scale_color_viridis() used by every panel below comes from viridis; the
# source notebook relied on it being attached by the interactive session.
library(viridis)

# Paths. REPAIR: the notebook used a hard-coded /scicore/... prefix and wrote
# figures with paste0(fpath, "name.pdf"), which depends on fpath carrying a
# trailing slash. Both are replaced by here::here()-derived paths and psave().
path_sim <- file.path(pp, "documents", "rPCRsim")
fpath    <- file.path(pp, "documents", "figures")
# Output naming (added when the figure documents were put on one convention).
# documents/figures holds the manuscript assets, one file per panel or panel
# range, named ratioPCR_figure_<figure><panel>. documents/figures/components
# holds the pieces that were assembled into a panel outside R, the diagnostic
# plots and the alternative views, under their working names.
cpath <- file.path(fpath, "components")
dir.create(fpath, recursive = TRUE, showWarnings = FALSE)
dir.create(cpath, recursive = TRUE, showWarnings = FALSE)

sim_scycle <- "Sim_Hetero_sCycle"

# Re-simulation switch, FALSE by default: OpenModelica and the rPCR_03 model
# tree are not available at render time. No Quarto `params:` block is used.
resimulate <- as.logical(Sys.getenv("RPCR_RESIMULATE", "FALSE"))
simModel   <- "rPCR_03"
simPath    <- file.path(Sys.getenv("HOME"), "openmodelica", simModel)

# Panel theme. REPAIR (session leakage): inherited from the interactive
# session in the source notebook.
theme_figure <- theme(
  axis.text.y = element_text(size = 5, face = "plain"),
  axis.text.x = element_text(size = 5),
  panel.grid.major = element_blank(),
  axis.text = element_text(size = 5),
  axis.title = element_text(face = "plain"),
  axis.title.y = element_text(vjust = 0),
  strip.placement = "outside",
  strip.text.y = element_text(vjust = 0, size = 8),
  plot.margin = margin(0, 0, 0, 0, "cm"))

# Standard kinetic parameter set, verbatim from the source notebook. The
# single-cycle design below is this set with cycle_max = 1 and a swept
# starting template amount.
stdPara <- data.frame(
  tolerance = 1e-8,
  cycle_max = 45,
  cycle_time = 45,
  startPrimer = 500 * 1e-9, # nM
  startTemplate = copytoconc(1e6), # copies
  startRatio = 0.5,
  ProbeA = 0 * 1e-9, # nM
  ProbeB = 0 * 1e-9, # nM
  Signal_A = 0,
  Signal_B = 0,
  Primer_k1 = 2e5,
  # literature: 5e5 /M s
  Primer_k2 = 1e-3,
  # literature: 1e-4 /s
  ProbeA_k1 = 3e5,
  ProbeA_k2 = 1e-3,
  ProbeB_k1 = 3e5,
  ProbeB_k2 = 1e-3,
  hetero_k1 = 1e6,
  hetero_k2 = 1e-4,
  rehybA_k1 = 1e6,
  rehybA_k2 = 1e-4,
  rehybB_k1 = 1e6,
  rehybB_k2 = 1e-4,
  Polymerase = 2e-7,
  RdsPoly_k1 = 1e6,
  RdsPoly_k2 = 1e-3,
  RiniPoly_k1 = 1e6,
  RiniPoly_k2 = 1e-3,
  RextPoly_k1 = 60,
  sFluoA = 50e6,
  sFluoB = 50e6,
  bFluoA = 1,
  bFluoB = 1,
  lFluoA = 0,
  lFluoB = 0,
  Aa_PB_k1 = 1e-20,
  Aa_PB_k2 = 1e20,
  Ba_PA_k1 = 1e-20,
  Ba_PA_k2 = 1e20
)

stdPara$stopTime <- stdPara$cycle_time * stdPara$cycle_max
stdPara$stepSize <- stdPara$cycle_time

results <- readRDS(file.path(path_sim, sim_scycle, "results", "results.rds"))
```

------------------------------------------------------------------------

# Panel a — Simulation design

Panel a states the design that panels b–d read out. A single PCR cycle is simulated over a two-dimensional grid: ten total template amounts spanning essentially zero to 95% of the primer concentration, and starting proportions from 0 to 1 in steps of 0.05. The primer concentration is reduced by the template amount at each grid point so that total nucleic acid, and hence the re-annealing competition, is held comparable across the sweep. The homoduplex rehybridisation rates of both templates are fixed at k₁_homo = 1×10⁶ M⁻¹s⁻¹ and the heteroduplex association rate is set to 1×10⁶, 6×10⁵ or 3×10⁵ M⁻¹s⁻¹ — equal to, moderately below, and well below the homoduplex rate.

From each simulated cycle five quantities are derived as a function of the starting proportion: the fraction of strands re-annealed as homoduplex, the fraction re-annealed as heteroduplex, their sum (total re-annealing), amplification efficiency (new strands relative to input strands) and absolute new amplicon yield. Efficiencies are formed per template (`effA`, `effB`) as well as for the total (`effT`), because the whole argument of this figure turns on the *difference* between the two.

```{r}
#| label: figS6-panel-a

title <- "Hetero_sCycle"
expName <- paste0("Sim_",title)
para <- stdPara
para$stepSize <- para$cycle_time
para$cycle_max<-1
para$stopTime <-para$cycle_time * para$cycle_ma

template <- seq(1e-18,para$startPrimer*0.95,length.out=10)
primer <- para$startPrimer-template
mpara <- cbind(para, rep(1,length(template)))
mpara$startTemplate <- template
mpara$startPrimer <- primer

nlist <- list(hetero_k1 = c(3e5,6e5,1e6),
              startRatio = seq(0.0,1,0.05))

if (resimulate) {
  results <- omPCRBatch(simModel,
                        simPath,
                        override=mpara,
                        nestedList=nlist,
                        logdir=expName,
                        project=title,
                        details = FALSE,
                        expName=expName)
}

results$sprop <- results$par_startRatio
results$prop <- results$ratio
results$sratio <- results$sprop/(1-results$sprop)
results$slratio <- log(results$sratio)
results$ratio <- results$prop/(1-results$prop)
results$lratio <- log(results$ratio)
results$sT <- results$par_startTemplate 
results$sA <- results$sT * results$sprop
results$sB <- results$sT * (1- results$sprop)
results$difA <- results$totalA -results$sA
results$difB <- results$totalB -results$sB
results$difT <- results$totalT -results$sT
results$effA <- results$difA/results$sA 
results$effB <- results$difB/results$sB
results$effT <- results$difT/results$sT

results <- results  %>%
  mutate(homo = AaAs.c +  BaBs.c +  AaAsPoly.c + BaBsPoly.c ,
         hetero = AaBs.c + AsBa.c + AaBsPoly.c + AsBaPoly.c,
         homoA = AaAs.c + AaAsPoly.c,
         homoB = BaBs.c + BaBsPoly.c)
```

`para$stopTime <- para$cycle_time * para$cycle_ma` is reproduced as written: `cycle_ma` is a typo for `cycle_max`, and because `para` is a data frame, partial matching resolves `para$cycle_ma` to `cycle_max`, so the line does what it appears to do. It is left unchanged rather than "corrected", since altering it cannot change the result and any edit risks changing it.

The design grid actually present in the stored results is reported below rather than restated from the legend.

```{r}
#| label: figS6-design-grid

design <- data.frame(
  quantity = c("total template amount sT (M)",
               "starting proportion",
               "heteroduplex association rate k1_hetero (1/M/s)",
               "homoduplex rehybridisation rate k1_homo (1/M/s)",
               "cycles simulated"),
  n_levels = c(length(unique(results$sT)),
               length(unique(results$sprop)),
               length(unique(results$par_hetero_k1)),
               1L,
               length(unique(results$cycle))),
  values = c(paste(format(range(results$sT), digits = 3), collapse = " to "),
             paste(format(range(results$sprop)), collapse = " to "),
             paste(format(sort(unique(results$par_hetero_k1)), scientific = TRUE),
                   collapse = ", "),
             format(stdPara$rehybA_k1, scientific = TRUE),
             paste(sort(unique(results$cycle)), collapse = ", "))
)
knitr::kable(design, row.names = FALSE,
             caption = paste("Single-cycle design grid read back from",
                             sim_scycle, "results.rds. The lowest template level",
                             "is 1e-18 M, i.e. effectively zero, and is excluded",
                             "from the per-template analyses of panel d."))
```

The figure itself shows this design as an annotated schematic; that drawing was prepared outside R. What is generated from code here is the design grid and the five derived quantities, which panels b and c plot.

------------------------------------------------------------------------

# Panel b — Equal association rates (k₁_hetero = 1×10⁶ M⁻¹s⁻¹)

With the heteroduplex rate equal to the homoduplex rate, re-annealing is blind to which strand it captures. Homoduplex formation is therefore *minimal* and heteroduplex formation *maximal* at equimolar input, simply because at 1:1 the probability that a given strand meets a non-complementary partner is highest. The two effects are exactly complementary, so their sum — total re-annealing — and with it amplification efficiency, is flat in the starting proportion. This is the control case: heteroduplexes form abundantly and yet no proportion-dependent efficiency difference, and hence no ratio distortion, arises. Efficiency does fall with increasing total template amount (colour), because re-annealing competes with primer annealing more strongly when more strands are present.

```{r}
#| label: figS6-panel-b
#| fig-cap: "Panel b. Equal association rates. Left to right: homoduplex re-annealing, heteroduplex re-annealing, total re-annealing, total amplification efficiency, and new amplicon yield, each against starting template proportion. Colour encodes the total template amount."
#| fig-width: 12
#| fig-height: 2.8

results <- results  %>%
  mutate(homo = AaAs.c +  BaBs.c +  AaAsPoly.c + BaBsPoly.c ,
         hetero = AaBs.c + AsBa.c + AaBsPoly.c + AsBaPoly.c)

sel <- results[results$par_hetero_k1==1e6,]

plist <- list()

plist[[1]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homo-2*difT)/sT*100,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of homoduplex (%)")

plist[[2]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=hetero/sT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of heteroduplex (%)")

plist[[4]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homo-2*difT+hetero)/sT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total re-annealing (%)")

plist[[5]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=effT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total efficiency (%)")

plist[[6]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=difT,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")+ xlab("proportion") + ylab("new amplicons (M)")

panel_b <- ggarrange(plotlist=plist[c(1,2,4,5,6)],ncol=5,nrow=1,common.legend=TRUE,legend="right")
panel_b
```

`plist[[3]]` is skipped in the source (it was the heteroduplex-share panel, commented out), and the arrangement selects elements 1, 2, 4, 5, 6. The gap in the index is preserved so that the element numbering matches the source notebook.

------------------------------------------------------------------------

# Panel c — Reduced heteroduplex association rate (k₁_hetero = 3×10⁵ M⁻¹s⁻¹)

When heteroduplexes form less readily than homoduplexes — the physically realistic case, since a heteroduplex carries the mismatch of the alternatively spliced segment — the compensation of panel b breaks down. Heteroduplex formation is still maximal at equimolar input, but it no longer replaces homoduplex formation one for one, so total re-annealing is *lowest* at 1:1 and amplification efficiency is correspondingly *highest* there. Efficiency now depends on the starting proportion.

The top row repeats the five total-template views of panel b at the reduced rate; the two rows below decompose the same simulations into template A and template B. The decomposition is what shows that the efficiency penalty is not shared equally: each template's heteroduplex burden is scaled by the *other* template's abundance, so the minority template loses proportionally less.

```{r}
#| label: figS6-panel-c-total
#| fig-cap: "Panel c, top rows. Total-template view at k1_hetero = 3e5: homoduplex, heteroduplex and total re-annealing, total efficiency and new amplicon yield against starting proportion, coloured by total template amount."
#| fig-width: 12
#| fig-height: 2.8

sel <- results[results$par_hetero_k1==3e5,]

plist <- list()

plist[[1]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homo-2*difT)/sT*100,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of homoduplex (%)")

plist[[2]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=hetero/sT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of heteroduplex (%)")

plist[[4]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homo-2*difT+hetero)/sT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total re-annealing (%)")

plist[[5]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=effT*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total efficiency (%)")

plist[[6]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=difT,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")+ xlab("proportion") + ylab("new amplicons (M)")

panel_c_total <- ggarrange(plotlist=plist[c(1,2,4,5,6)],ncol=5,nrow=1,common.legend=TRUE,legend="right")
panel_c_total
```

```{r}
#| label: figS6-panel-c-templateA
#| fig-cap: "Panel c, template A. Homoduplex re-annealing, heteroduplex re-annealing, total re-annealing and amplification efficiency of template A alone, at k1_hetero = 3e5, against starting proportion."
#| fig-width: 10
#| fig-height: 2.8

### oonly template A

sel <- results[results$par_hetero_k1==3e5,]

plist <- list()

plist[[1]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homoA-2*difA)/sA*100,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of homoduplex (%)")

plist[[2]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(hetero/2)/sA*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of heteroduplex (%)")

plist[[3]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=hetero/(homoA*2 + hetero),group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(0,1) + xlab("proportion") + ylab("template A heteroduplex (%)")

plist[[4]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homoA-2*difA+hetero/2)/sA*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total re-annealing (%)")

plist[[5]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=effA*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("efficiency (%)")

plist[[6]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=difA,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")+ xlab("proportion") + ylab("new amplicons (M)")

panel_c_A <- ggarrange(plotlist=plist[c(1,2,4,5)],ncol=4,nrow=1,common.legend=TRUE,legend="right")
panel_c_A
```

```{r}
#| label: figS6-panel-c-templateB
#| fig-cap: "Panel c, template B. The same four views for template B. Comparing with template A shows the efficiency asymmetry: the two curves are mirror images about the equimolar point."
#| fig-width: 10
#| fig-height: 2.8

### only template B

sel <- results[results$par_hetero_k1==3e5,]

plist <- list()

plist[[1]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homoB-2*difB)/sB*100,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of homoduplex (%)")

plist[[2]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(hetero/2)/sB*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("re-annealing of heteroduplex (%)")

plist[[3]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=hetero/(homoA*2 + hetero),group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(0,1) + xlab("proportion") + ylab("template A heteroduplex (%)")

plist[[4]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=(homoB-2*difB+hetero/2)/sB*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("total re-annealing (%)")

plist[[5]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=effB*100,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal") + ylim(-1,100) + xlab("proportion") + ylab("efficiency (%)")

plist[[6]] <- ggplot() + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(data=sel,aes(x=par_startRatio,y=difB,group=interaction(par_hetero_k1,par_startTemplate),col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")+ xlab("proportion") + ylab("new amplicons (M)")

panel_c_B <- ggarrange(plotlist=plist[c(1,2,4,5)],ncol=4,nrow=1,common.legend=TRUE,legend="right")
panel_c_B
```

The template-B block reuses `plist[[3]]` from the template-A block unchanged (`hetero/(homoA*2 + hetero)`, labelled "template A heteroduplex"). That element is not part of the assembled panel — the arrangement selects 1, 2, 4, 5 — so the copy has no effect on the figure and is left as written rather than silently retargeted to `homoB`.

------------------------------------------------------------------------

# Panel d — Proportion-dependent efficiency differences drive equimolar drift

Panel d is the conclusion of the argument. Its three columns are the three heteroduplex association rates, 1×10⁶, 6×10⁵ and 3×10⁵ M⁻¹s⁻¹. The top row plots the two per-template efficiencies against one another: at equal rates the locus collapses onto a single point per template amount (the efficiencies are identical at every proportion), and it opens up as the heteroduplex rate falls. The middle row shows the efficiency ratio effA/effB on a logarithmic axis, which deviates from 1 in a proportion-dependent, sign-changing way. The bottom row is the consequence for measurement: the residual between the amplicon proportion after one cycle and the input proportion, which traces the S-shaped bias profile of Figure 2G in miniature. It crosses zero at the equimolar point, where both templates face identical re-annealing competition.

```{r}
#| label: figS6-panel-d
#| fig-cap: "Panel d. Columns: k1_hetero = 1e6, 6e5, 3e5. Top row: template A efficiency against template B efficiency, coloured by starting proportion. Middle row: efficiency ratio effA/effB against starting proportion (log axis), coloured by total template amount. Bottom row: residual (amplicon proportion minus template proportion) against template proportion."
#| fig-width: 11
#| fig-height: 8

plist <- list()

sel <- results[results$par_hetero_k1==1e6,]

plist[[1]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=effA/effB,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")   + xlab("proportion") + ylab(" efficiency ratio (template A/template B)") + scale_y_log10(limits = c(0.5,2))

plist[[2]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=effA,y=effB,group=sprop,col=sprop)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
  labs(x="template A efficiency", y="template B efficiency", color = "proportion")

plist[[7]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=sprop,y=prop-sprop,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
  labs(x="template proportion", y="residuals (amplicon proportion - template proportion", color = "proportion")+ ylim(-0.015,+0.015)

sel <- results[results$par_hetero_k1==6e5,]

plist[[3]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=effA/effB,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")   + xlab("proportion") + ylab(" efficiency ratio (template A/template B)") + scale_y_log10(limits = c(0.5,2))

plist[[4]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=effA,y=effB,group=sprop,col=sprop)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
  labs(x="template A efficiency", y="template B efficiency", color = "proportion")

plist[[8]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=sprop,y=prop-sprop,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
labs(x="template proportion", y="residuals (amplicon proportion - template proportion", color = "proportion")+ ylim(-0.015,+0.015)

sel <- results[results$par_hetero_k1==3e5,]

plist[[5]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=par_startRatio,y=effA/effB,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")   + xlab("proportion") + ylab(" efficiency ratio (template A/template B)") + scale_y_log10(limits = c(0.5,2))

plist[[6]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=effA,y=effB,group=sprop,col=sprop)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
  labs(x="template A efficiency", y="template B efficiency", color = "proportion")

plist[[9]] <- ggplot(data=sel) + theme_rpcr() + theme_figure + scale_color_viridis(discrete=FALSE) +
  geom_line(aes(x=sprop,y=prop-sprop,group=sT,col=sT)) +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.box = "horizontal")  +
labs(x="template proportion", y="residuals (amplicon proportion - template proportion", color = "proportion") + ylim(-0.015,+0.015)

# REPAIR: in the source notebook this arrangement was printed but never
# assigned, and the ggsave() call that followed wrote the *previous* chunk's
# `plot` object (the template-B arrangement) under the panel-d file name. The
# arrangement is assigned here so that the exported file is the panel that was
# built.
panel_d <- ggarrange(
  ggarrange(plotlist=plist[c(2,4,6)],ncol=3,nrow=1,common.legend=TRUE,legend="right"),
  ggarrange(plotlist=plist[c(1,3,5)],ncol=3,nrow=1,common.legend=TRUE,legend="right"),
  ggarrange(plotlist=plist[c(7,8,9)],ncol=3,nrow=1,common.legend=TRUE,legend="right"),
  nrow=3)

panel_d
```

Two further edits to this block are worth recording. The `labs(...,, color = ...)` calls in the source contain a doubled comma, which R parses as an empty argument and `labs()` tolerates; the duplicate comma is removed because it carries no information. And the exploratory material that followed in the notebook — `plot_ly()` efficiency surfaces and a series of `nlsLM()` fits of candidate analytical drift functions (`propdrift`, `drift`, `pdrift`, `rdrift`, `sdrift`, `odrift`, `ndrift`) — is not carried over: none of it produces a panel of this figure, it requires `plotly` and `minpack.lm` at render time, and several of those fits are superseded by the analytical calibration model of Figure 3.

The claim that the residual crosses zero at equimolar input, and that the efficiency ratio departs from 1 only when the heteroduplex rate is reduced, is checked here against the data rather than asserted.

```{r}
#| label: figS6-drift-check

check <- do.call(rbind, lapply(sort(unique(results$par_hetero_k1)), function(k) {
  s <- results[results$par_hetero_k1 == k & results$sT > min(results$sT), ]
  eq <- s[abs(s$sprop - 0.5) < 1e-9, ]
  data.frame(
    k1_hetero = format(k, scientific = TRUE),
    max_abs_efficiency_ratio_dev = max(abs(s$effA / s$effB - 1), na.rm = TRUE),
    max_abs_residual = max(abs(s$prop - s$sprop), na.rm = TRUE),
    max_abs_residual_at_equimolar = max(abs(eq$prop - eq$sprop), na.rm = TRUE)
  )
}))
knitr::kable(check, digits = 6, row.names = FALSE,
             caption = paste("Per heteroduplex association rate, over all template",
                             "amounts above the near-zero level and all starting",
                             "proportions: largest departure of the per-template",
                             "efficiency ratio from 1, largest residual between",
                             "amplicon and template proportion, and the residual at",
                             "the equimolar point."))
```

------------------------------------------------------------------------

# Panel e — Schematic of the inhibition mechanism

Panel e is a **schematic** and has no R source in this repository. It summarises the mechanism that panels b–d establish numerically: re-annealing of amplicon strands inhibits amplification because a duplex is not a substrate for primer extension, but heteroduplexes inhibit *less* than homoduplexes when their association rate is lower. The minority template holds a proportionally larger share of its strands in heteroduplexes and therefore suffers less inhibition, giving it the higher effective efficiency. The result is a systematic pull of the measured ratio toward equimolarity whose magnitude grows as k₁_hetero falls relative to k₁_homo.

The drawing was prepared outside R and is not generated from code here.

------------------------------------------------------------------------

# Save figure

```{r}
#| label: save-figures

# REPAIR: the source notebook wrote these with
# ggsave(paste0(fpath, "name.pdf"), ...), which depends on fpath carrying a
# trailing slash and bypasses the project's save wrapper. All writes now go
# through psave() into fpath. Sizes are those of the source calls; the panel-d
# export used the same 6 x 3 cm as the single-row strips, which is too small
# for a 3 x 3 arrangement, so it is written at 18 x 12 cm.
psave("ratioPCR_figure_S6b",          panel_b,       width = 6,  height = 3,  dpi = 600, units = "cm", device = "pdf")
psave("ratioPCR_figure_S6c",          panel_c_total, width = 6,  height = 3,  dpi = 600, units = "cm", device = "pdf")
psave("Hetero_sCycle_3e5_templateA",  panel_c_A, path = cpath, width = 6,  height = 3,  dpi = 600, units = "cm", device = "pdf")
psave("Hetero_sCycle_3e5_templateB",  panel_c_B, path = cpath, width = 6,  height = 3,  dpi = 600, units = "cm", device = "pdf")
psave("ratioPCR_figure_S6d",          panel_d,       width = 18, height = 12, dpi = 600, units = "cm", device = "pdf")
```
