---
title: "ratioPCR Figure S8 — Pearson correlation between rPCR and dPCR inclusion proportion measurements across all three assays"
format:
  html:
    embed-resources: true
    code-background: true
    code-fold: true
    code-tools: true
    toc: true
    toc-depth: 3
    number-sections: true
editor: visual
execute:
  warning: false
  message: false
---

<!-- ============================================================= -->
<!-- ratioPCR_Figure_S8.qmd                                         -->
<!-- Self-contained rebuild of Supplementary Figure S8              -->
<!-- (ISCIENCE-D-26-02393_R2).                                      -->
<!-- ============================================================= -->

# Figure legend

> **Supplementary Figure S8: Pearson correlation between rPCR and dPCR
> inclusion proportion measurements across all three assays**
>
> Correlation between calibrated rPCR inclusion proportions and dPCR-derived
> inclusion proportions for KIF13A, NUMA1, and SYNE1. Each point represents the
> mean proportion across replicates for one reference mixture (n = 11 reference
> mixtures spanning 0–100% inclusion). Error bars show standard error of the
> mean. Solid lines show linear regression fits; shaded areas indicate 95%
> confidence intervals. Pearson correlation coefficients and p-values are shown
> in each panel. KIF13A and SYNE1 show near-perfect correlation (R = 1, p <
> 2.5×10⁻¹³), while NUMA1 shows slightly reduced correlation (R = 0.97, p =
> 4.2×10⁻⁷), consistent with its greater measurement scatter (RSE = 0.071,
> Figure 5C).

------------------------------------------------------------------------

# Setup

This figure and main Figure 5C are produced by the same script:
Figure 5C is its `panelE`
(theoretical against measured proportion per method) and Figure S8 is its
`panelSup` (rPCR against dPCR, `plot3_K`, `plot3_N`, `plot3_S`). The data
preparation is therefore identical in the two documents and is repeated here so
that this document is self-contained.


```{r}
#| label: setup

pp <- here::here()
source(file.path(pp, "documents", "scripts", "Figure_settings.R"))

# Output naming (added when the figure documents were put on one convention).
# documents/figures holds the manuscript assets, one file per panel or panel
# range, named ratioPCR_figure_<figure><panel>. documents/figures/components
# holds the pieces that were assembled into a panel outside R, the diagnostic
# plots and the alternative views, under their working names.
cpath <- file.path(fpath, "components")
dir.create(fpath, recursive = TRUE, showWarnings = FALSE)
dir.create(cpath, recursive = TRUE, showWarnings = FALSE)

library(dplyr)

ppath <- file.path(pp, "documents", "extdata", "dPCR")

# KIF13A: rPCR run and dPCR partition summary of the same reference series
import_K <- rpcr_import(file.path(ppath, "rPCR_20250916_JK KIF13A dPCR comparison_20250918_125346.xlsx"))
dPCR_K   <- read.csv(file.path(ppath, "rPCR_20250916 KIF13A_analysis_18_09_2025_17_22_21_UTC+02_00.csv"), skip = 1)

# NUMA1
import_N <- rpcr_import(file.path(pp, "documents", "extdata", "rPCR_NUMA1_compare.xlsx"))
dPCR_N   <- read.csv(file.path(pp, "documents", "extdata", "dPCR_NUMA1_compare.csv"), sep = ";", header = FALSE)

# SYNE1
import_S <- rpcr_import(file.path(pp, "documents", "extdata", "rPCR_SYNE1_compare.xlsx"))
dPCR_S   <- read.csv(file.path(ppath, "dPCR_SYNE1_compare.csv"), skip = 1)

# Collector for the correlation coefficients reported by the source script
cor_all <- data.frame()
```

------------------------------------------------------------------------

# Panel — KIF13A

Each of the 11 reference mixtures is measured by both methods on the same
material. The point plotted is the replicate mean per mixture, with the SEM in
both directions, so that the regression is fitted on the quantity the two
methods are meant to agree on rather than on the replicate scatter. The
identity line is drawn as a dotted reference; agreement with it is what
distinguishes correlation from accuracy, which is reported separately as the
RSE in Figure 5C.

The rPCR side uses the `lbKIF13A` reference series of the dPCR-comparison
plate. A first calibration of the `laKIF13A` series in the source script is
overwritten two statements later without being used, and has been dropped as
dead code.

```{r}
#| label: figS8-kif13a

target <- "KIF13A"

import <- import_K
dPCR   <- dPCR_K

### correlation
df2 <- dPCR[dPCR$"Target..Name." == "Incl" & dPCR$"Sample.NTC.Control" != "NTC", c("Well", "Edited.fraction")]
df2$row <- substring(df2$Well, 1, 1)
df2$col <- as.numeric(substring(df2$Well, 2))
seriesA <- c(0, 0.05, 0.15, 0.25, 0.35, 0.5, 0.65, 0.75, 0.85, 0.95, 1)
seriesB <- seq(0, 1, 0.1)

df2$tprop <- c(rep(seriesA, 4), rep(seriesB, 4))
df2$cprop <- as.numeric(gsub("%", "", df2$"Edited.fraction")) / 100
df2$delta <- df2$cprop - df2$tprop
df2 <- df2[substr(df2$Well, 1, 1) %in% c("E", "F", "G", "H"), ]

res <- rpcr_calibrate(import[import$target == "lbKIF13A", ], qual = "Cal", output = "details")
df1 <- res$summary
df1 <- rpcr_calibDetect(df1[df1$sample != "NTC", ])
df1$delta <- df1$cprop - df1$tprop

df1$method <- "rPCR"
df2$method = "dPCR"
colnames(df2)[colnames(df2) == "Well"] <- "well"

df1_mean <- df1 %>% group_by(tprop) %>% summarize(mean_cprop1 = mean(cprop, na.rm = TRUE))
df2_mean <- df2 %>% group_by(tprop) %>% summarize(mean_cprop2 = mean(cprop, na.rm = TRUE))

merged_means <- inner_join(df1_mean, df2_mean, by = "tprop")
cor_all <- rbind(cor_all, data.frame(target = "KIF13A", type = "RvsD", R = cor(merged_means$mean_cprop1, merged_means$mean_cprop2)))
cor_all <- rbind(cor_all, data.frame(target = "KIF13A", type = "RvsT", R = cor(df1$tprop, df1$cprop)))
cor_all <- rbind(cor_all, data.frame(target = "KIF13A", type = "DvsT", R = cor(df2$tprop, df2$cprop)))

stats1 <- df1 %>% group_by(tprop) %>%
  summarize(m1 = mean(cprop), sd1 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))
stats2 <- df2 %>% group_by(tprop) %>%
  summarize(m2 = mean(cprop), sd2 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))

plot_data <- inner_join(stats1, stats2, by = "tprop")
plot_data_K <- plot_data

plot3_K <- ggplot(plot_data, aes(x = m1, y = m2)) +
  geom_line(data = data.frame(x = c(0, 1), y = c(0, 1)), aes(x = x, y = y), linetype = "dotted", linewidth = 0.5, col = "grey") +
  # Fehlerbalken für beide Achsen
  geom_errorbar(aes(ymin = m2 - sd2, ymax = m2 + sd2), width = 0.02, color = "gray") +
  geom_errorbar(aes(xmin = m1 - sd1, xmax = m1 + sd1), height = 0.02, color = "gray") +
  # Mittelpunkte
  geom_point(color = "black", size = 0.5, alpha = 0.2) +
  # Regressionslinie
  geom_smooth(method = "lm", color = "grey50", se = TRUE) +
  # Korrelationskoeffizient und p-Wert automatisch einblenden
  stat_cor(method = "pearson", label.x = 0.1, label.y = 0.9) +
  labs(x = "proportion (rPCR)", y = "proportion (dPCR)") +
  theme_rpcr()
```

```{r}
#| label: figS8-kif13a-plot
#| fig-width: 2.4
#| fig-height: 2.4

plot3_K
```

------------------------------------------------------------------------

# Panel — NUMA1

```{r}
#| label: figS8-numa1

target <- "NUMA1"

import <- import_N
import <- import[import$sample != "NTC" & import$target == "cNUMA1", ]
calib <- rpcr_calibrate(import, qual = "Cal2_")
res <- rpcr_batchFit(import, calib = calib)
res <- rpcr_calibDetect(res, qual = "Cal2_")

df1 <- res
df1 <- df1[grep("Cal2", df1$sample), ]
df1$delta <- df1$cprop - df1$tprop

dPCR <- dPCR_N
tprop <- c(0, 0.05, 0.15, 0.25, 0.35, 0.5, 0.65, 0.75, 0.85, 0.95, 1)
tprop <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)
colnames(dPCR) <- c("well", "dil", tprop)

df2 <- dPCR[dPCR$dil == "dilB", ] %>%
  pivot_longer(
    cols = as.character(tprop),
    names_to = "tprop",
    values_to = "cprop"
  )
df2$tprop <- as.numeric(df2$tprop)
df2$delta <- df2$cprop - df2$tprop

df1$method <- "rPCR"
df2$method = "dPCR"

df1_mean <- df1 %>% group_by(tprop) %>% summarize(mean_cprop1 = mean(cprop, na.rm = TRUE))
df2_mean <- df2 %>% group_by(tprop) %>% summarize(mean_cprop2 = mean(cprop, na.rm = TRUE))

merged_means <- inner_join(df1_mean, df2_mean, by = "tprop")
cor_all <- rbind(cor_all, data.frame(target = "NUMA1", type = "RvsD", R = cor(merged_means$mean_cprop1, merged_means$mean_cprop2)))
cor_all <- rbind(cor_all, data.frame(target = "NUMA1", type = "RvsT", R = cor(df1$tprop, df1$cprop)))
cor_all <- rbind(cor_all, data.frame(target = "NUMA1", type = "DvsT", R = cor(df2$tprop, df2$cprop)))

stats1 <- df1 %>% group_by(tprop) %>%
  summarize(m1 = mean(cprop), sd1 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))
stats2 <- df2 %>% group_by(tprop) %>%
  summarize(m2 = mean(cprop), sd2 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))

plot_data <- inner_join(stats1, stats2, by = "tprop")
plot_data_N <- plot_data

plot3_N <- ggplot(plot_data, aes(x = m1, y = m2)) +
  geom_line(data = data.frame(x = c(0, 1), y = c(0, 1)), aes(x = x, y = y), linetype = "dotted", linewidth = 0.5, col = "grey") +
  geom_errorbar(aes(ymin = m2 - sd2, ymax = m2 + sd2), width = 0.02, color = "gray") +
  geom_errorbar(aes(xmin = m1 - sd1, xmax = m1 + sd1), height = 0.02, color = "gray") +
  geom_point(color = "black", size = 0.5, alpha = 0.2) +
  geom_smooth(method = "lm", color = "grey50", se = TRUE) +
  stat_cor(method = "pearson", label.x = 0.1, label.y = 0.9) +
  labs(x = "proportion (rPCR)", y = "proportion (dPCR)") +
  theme_rpcr()
```

```{r}
#| label: figS8-numa1-plot
#| fig-width: 2.4
#| fig-height: 2.4

plot3_N
```

------------------------------------------------------------------------

# Panel — SYNE1

The SYNE1 dPCR export reports the *exclusion* target, so the inclusion
proportion is `1 − edited fraction`. Calibration of the rPCR side uses four
standards (`Cal10`, `Cal40`, `Cal60`, `Cal90`) rather than the full series.

```{r}
#| label: figS8-syne1

target <- "SYNE1"

import <- import_S
dPCR   <- dPCR_S

refs <-  c("Cal10", "Cal40", "Cal60", "Cal90")
calib <- rpcr_calibrate(import[import$sample %in% refs, ])
res <- rpcr_batchFit(import, calib = calib)
res <- rpcr_calibDetect(res)

### correlation
df2 <- dPCR[grepl("Cal", dPCR$Sample.NTC.Control) & grepl("excl", dPCR$"Target..Name.") & dPCR$"Sample.NTC.Control" != "NTC", c("Well", "Edited.fraction")]
df2$row <- substring(df2$Well, 1, 1)
df2$col <- as.numeric(substring(df2$Well, 2))

df2$tprop <- rep(seq(0, 1, 0.1), 4)[1:43]
df2$cprop <- 1 - as.numeric(gsub("%", "", df2$"Edited.fraction")) / 100
df2$delta <- df2$cprop - df2$tprop
df2 <- na.omit(df2)

df1 <- res
df1 <- df1[grep("Cal", df1$sample), ]
df1$delta <- df1$cprop - df1$tprop

df1$method <- "rPCR"
df2$method = "dPCR"
colnames(df2)[colnames(df2) == "Well"] <- "well"

df1_mean <- df1 %>% group_by(tprop) %>% summarize(mean_cprop1 = mean(cprop, na.rm = TRUE))
df2_mean <- df2 %>% group_by(tprop) %>% summarize(mean_cprop2 = mean(cprop, na.rm = TRUE))

merged_means <- inner_join(df1_mean, df2_mean, by = "tprop")

cor_all <- rbind(cor_all, data.frame(target = "SYNE1", type = "RvsD", R = cor(merged_means$mean_cprop1, merged_means$mean_cprop2)))
cor_all <- rbind(cor_all, data.frame(target = "SYNE1", type = "RvsT", R = cor(df1$tprop, df1$cprop)))
cor_all <- rbind(cor_all, data.frame(target = "SYNE1", type = "DvsT", R = cor(df2$tprop, df2$cprop)))

stats1 <- df1 %>% group_by(tprop) %>%
  summarize(m1 = mean(cprop), sd1 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))
stats2 <- df2 %>% group_by(tprop) %>%
  summarize(m2 = mean(cprop), sd2 = sd(cprop) / sqrt(n())) %>%
  mutate(tprop = round(tprop, 2))

plot_data <- inner_join(stats1, stats2, by = "tprop")
plot_data_S <- plot_data

plot3_S <- ggplot(plot_data, aes(x = m1, y = m2)) +
  geom_line(data = data.frame(x = c(0, 1), y = c(0, 1)), aes(x = x, y = y), linetype = "dotted", linewidth = 0.5, col = "grey") +
  # Fehlerbalken für beide Achsen
  geom_errorbar(aes(ymin = m2 - sd2, ymax = m2 + sd2), width = 0.02, color = "gray") +
  geom_errorbar(aes(xmin = m1 - sd1, xmax = m1 + sd1), height = 0.02, color = "gray") +
  # Mittelpunkte
  geom_point(color = "black", size = 0.5, alpha = 0.2) +
  # Regressionslinie
  geom_smooth(method = "lm", color = "grey50", se = TRUE) +
  # Korrelationskoeffizient und p-Wert automatisch einblenden
  stat_cor(method = "pearson", label.x = 0.1, label.y = 0.9) +
  labs(x = "proportion (rPCR)", y = "proportion (dPCR)") +
  theme_rpcr()
```

```{r}
#| label: figS8-syne1-plot
#| fig-width: 2.4
#| fig-height: 2.4

plot3_S
```

------------------------------------------------------------------------

# Correlation coefficients

`stat_cor()` annotates each panel with the Pearson statistics it computes from
the same `plot_data`. The values are recomputed here explicitly with
`cor.test()` so that the numbers quoted in the manuscript can be checked rather
than restated, and any disagreement reported.

```{r}
#| label: figS8-correlations

pearson_row <- function(target, d) {
  ct <- cor.test(d$m1, d$m2, method = "pearson")
  data.frame(target = target,
             n = nrow(d),
             R = unname(ct$estimate),
             p_value = ct$p.value)
}

cor_tbl <- rbind(
  pearson_row("KIF13A", plot_data_K),
  pearson_row("NUMA1",  plot_data_N),
  pearson_row("SYNE1",  plot_data_S)
)

# Values as reported in the Results text and legend of ISCIENCE-D-26-02393_R2.
cor_manuscript <- data.frame(
  target        = c("KIF13A", "NUMA1", "SYNE1"),
  R_manuscript  = c(1, 0.97, 1),
  p_manuscript  = c(2.4e-13, 4.2e-7, 2.5e-14)
)

cor_check <- merge(cor_tbl, cor_manuscript, by = "target")
cor_check$R_agrees <- abs(round(cor_check$R, 2) - cor_check$R_manuscript) < 0.005
cor_check$p_ratio  <- cor_check$p_value / cor_check$p_manuscript

knitr::kable(cor_check, digits = c(0, 0, 4, 12, 2, 12, 0, 3), row.names = FALSE,
             caption = "Recomputed Pearson correlation against the values quoted in the manuscript")

if (all(cor_check$R_agrees)) {
  cat("All three recomputed correlation coefficients reproduce the manuscript values to two decimals.\n")
} else {
  cat("DISAGREEMENT with the manuscript for:\n")
  print(cor_check[!cor_check$R_agrees, c("target", "R", "R_manuscript", "p_value", "p_manuscript")])
}

# The correlation collector assembled by the source script, for reference:
#   RvsD = rPCR vs dPCR (this figure), RvsT / DvsT = each method vs theoretical
cor_all
```

------------------------------------------------------------------------

# Assembled figure

```{r}
#| label: figS8-assembled
#| fig-width: 2.4
#| fig-height: 7

panelSup <- (plot3_K ) +
  (plot3_N ) +
  (plot3_S ) +
  plot_layout(
    nrow = 3,           # one panel per assay, stacked
    heights = c(1, 1, 1)
  )

panelSup
```

------------------------------------------------------------------------

# Save figure

The output file name written by the source script
(`Figure Supp (dPCR_compare).pdf`) is replaced by `ratioPCR_figure_S8.pdf`,
so that this document writes
the existing file in `documents/figures` rather than accumulating beside it.

```{r}
#| label: save-figures

dir.create(fpath, recursive = TRUE, showWarnings = FALSE)

psave("ratioPCR_figure_S8", panelSup, width = 5, height = 15, dpi = 600)

cat("Figure S8 output written to:", fpath, "\n")
```

------------------------------------------------------------------------

# Provenance and repairs

| Element | Source | Repairs applied |
|----|----|----|
| KIF13A, NUMA1, SYNE1 panels | `plot3_K` / `plot3_N` / `plot3_S` | Data imports hoisted to the setup chunk; the dead first KIF13A calibration block (`import0`, `laKIF13A`) removed; the `plot1_*` / `plot2_*` panels and the RSE summary belong to main Figure 5C and are built in `ratioPCR_Figure_5.qmd`. Plotting code unchanged, German inline comments included. |
| Assembled figure | `panelSup` → `ratioPCR_figure_S8.pdf` | Write routed through `psave()` in the save section. |

Known upstream quirk kept unchanged: the horizontal error bars use
`geom_errorbar(aes(xmin = , xmax = ), height = )`. `height` is not a
`geom_errorbar` parameter and is ignored with a warning (suppressed by the
document's execute options); the bars themselves render.

------------------------------------------------------------------------

# Manuscript figure

The files this document writes into `documents/figures`, re-printed here at
the size they are written at, so that the rendered document shows the panels
as they appear in the manuscript rather than only the analysis-sized views
above. The final chunk checks that every file is on disk.

**`ratioPCR_figure_S8.pdf`** — the whole figure, 5 × 15 cm.

```{r}
#| label: manuscript-figure-S8
#| echo: false
#| fig-width: 1.97
#| fig-height: 5.91

panelSup
```

```{r}
#| label: manuscript-figure-files
#| echo: false

written <- c("ratioPCR_figure_S8.pdf")
data.frame(file = written,
           exists = file.exists(file.path(fpath, written)),
           row.names = NULL)
```

# Reproducibility

```{r}
#| label: session-info
#| echo: true

if (requireNamespace("sessioninfo", quietly = TRUE)) {
  sessioninfo::session_info()
} else {
  print(sessionInfo())
}

if (requireNamespace("renv", quietly = TRUE)) {
  try(print(renv::status()), silent = TRUE)
}
```
