---
title: "ratioPCR Figure S7 — Ratio calculation using single cycles or multi-cycle slope fitting"
format:
  html:
    embed-resources: true
    code-background: true
    code-fold: true
    code-tools: true
    toc: true
    toc-depth: 3
    number-sections: true
editor: visual
execute:
  warning: false
  message: false
---

<<<BLOCK
<!-- ============================================================= -->
<!-- ratioPCR_Figure_S7.qmd                                         -->
<!-- Self-contained rebuild of Supplementary Figure S7.             -->
<!-- Chunk bodies transplanted; repairs listed at the end.          -->
<!-- ============================================================= -->
BLOCK>>>

# Figure legend

The supplementary legends are complete in the supplied legend file and are
quoted verbatim.

> Supplementary Figure S7: Ratio calculation using single cycles or
> multi-cycle slope fitting
>
> (A) Single-cycle ratio estimation using ΔRn and derRn signals. Uncalibrated
> and calibrated apparent inclusion proportions are shown for each cycle from
> C₋₁₀ to C₊₁₅ relative to Cᵢ for KIF13A, NUMA1, and SYNE1 using
> baseline-subtracted fluorescence (ΔRn, left four columns) and first-derivative
> signals (derRn, right four columns). Horizontal dashed lines indicate
> theoretical input proportions. Equivalent results for raw Rn signals are shown
> in Figure 4B.
>
> (B) Residuals between calibrated proportions and theoretical values as a
> function of relative cycle position, shown separately for Rn, ΔRn, and derRn
> signals for all three assays. Upper rows show the full deviation range; lower
> rows show violin plots of the deviation distribution across the cycle range
> C₋₅ to C₊₁₀, illustrating the dependence of accuracy and precision on both
> cycle selection and signal type. RSE comparison across signal types is shown
> in Figure 4C.
>
> (C) RSE heatmaps for the cycle-range slope fitting method. Each heatmap shows
> RSE values as a function of start and end cycle position relative to Cᵢ for
> KIF13A, NUMA1, SYNE1, and the mean across assays. Yellow regions indicate low
> RSE (optimal accuracy); purple regions indicate calibration failure. For each
> assay, an optimal window of cycle positions can be identified in which
> multi-cycle slope fitting achieves minimum RSE. The experimental design of the
> cycle-range method is illustrated in Figure 4D.
>
> (D) RSE heatmaps for the derivative-threshold slope fitting method. Each
> heatmap shows RSE values as a function of threshold 1 (x-axis, applied before
> Cᵢ) and threshold 2 (y-axis, applied after Cᵢ) for KIF13A, NUMA1, SYNE1, and
> the mean across assays. Warm colours indicate optimal threshold combinations;
> purple regions indicate calibration failure. The experimental design of the
> threshold method is illustrated in Figure 4E. Performance comparison of
> single-cycle and multi-cycle approaches is shown in Figure 4F.

| Panel | Content |
|-------|---------|
| A | Single-cycle ψₐ and ψ_c per cycle, ΔRn and derRn |
| B | Residuals by cycle position, violins over the central window |
| C | RSE heatmaps, cycle-range method |
| D | RSE heatmaps, derivative-threshold method |

# Setup

Panels A and B rest on the single-cycle sweep across relative cycle positions;
panels C and D on the two multi-cycle parameter sweeps. All three are expensive,
so all three respect the recompute switch:

```
RPCR_RECALCULATE=FALSE quarto render ratioPCR_Figure_S7.qmd   # read caches
RPCR_RECALCULATE=TRUE  quarto render ratioPCR_Figure_S7.qmd   # recompute (default)
```

```{r}
#| label: setup
#| echo: true

pp <- here::here()
source(file.path(pp, "documents", "scripts", "Figure_settings.R"))

# Output naming (added when the figure documents were put on one convention).
# documents/figures holds the manuscript assets, one file per panel or panel
# range, named ratioPCR_figure_<figure><panel>. documents/figures/components
# holds the pieces that were assembled into a panel outside R, the diagnostic
# plots and the alternative views, under their working names.
cpath <- file.path(fpath, "components")
dir.create(fpath, recursive = TRUE, showWarnings = FALSE)
dir.create(cpath, recursive = TRUE, showWarnings = FALSE)

library(ggh4x)
library(viridis)

# Render-time recompute switch (a plain variable, not Quarto's reserved `params`).
recalculate <- as.logical(Sys.getenv("RPCR_RECALCULATE", "TRUE"))

calibrationFile <- file.path(pp, "documents", "extdata", "20191120AS calibration.xlsx")
resultsCache    <- file.path(pp, "documents", "extdata", "results 2026-04-28 .rds")

# importing data
import <- rpcr_wrap(rpcr_readTF(calibrationFile))
import <- import[!import$sample == "neg. ctrl", ]
import <- rpcr_calibDetect(import, qual = "")

# `res` holds the sweep results keyed by analysis name. Without recompute it
# comes from the cache.
if (!recalculate && file.exists(resultsCache)) {
  res <- readRDS(resultsCache)
}

c(cache_present = file.exists(resultsCache),
  cached_analyses = paste(names(res), collapse = "; "))
```

# Single-cycle sweep across relative cycle positions

The same sweep that underlies Figure 4B and 4C: the full calibration is run
independently at each relative cycle from C₋₁₀ to C₊₁₅, for Rn and ΔRn and for
their untransformed (`0D`) and first-derivative (`1D`) forms, by collapsing the
fitting window onto a single cycle (`rtype = "ip"`, `roff = c(cycle, cycle)`).
Where Figure 4 shows the Rn results, panels A and B of this figure show the
ΔRn and derRn results and the residual structure of all three signal types.

The cache is shared with `ratioPCR_Figure_4.qmd`, so the sweep is computed once
per recompute cycle regardless of which of the two documents renders first.

```{r}
#| label: single-cycle-sweep
#| echo: true

targets <- unique(import$target)
datatypes <- c("Rn","deltaRn")
derivatives <- c("0D","1D")
crange <- c(-10:15)

singleCycleCache <- file.path(pp, "documents", "extdata", "single cycle results.rds")

if (!recalculate && file.exists(singleCycleCache)) {

  singlecycle  <- readRDS(singleCycleCache)
  results_data <- singlecycle$results_data
  results_qc   <- singlecycle$results_qc

} else {

  details <- setNames(lapply(datatypes, function(datatype)
    setNames(lapply(derivatives, function(derivative)
      setNames(lapply(crange,function(cycle)
        rpcr_calibrate(import,
                       qual = "",
                       output = "details",
                       signal = datatype,
                       derivative = if(derivative=="1D") TRUE else FALSE,
                       rtype = "ip", # method to define range
                       roff = c(cycle,cycle), # offset from range defined by rtype
                       model = "single" # fitting method: either linear or 2nd degree polynomial (poly)
        )
      ), paste("c ",crange))
    ), derivatives)
  ), datatypes)


  results_data <-   do.call(rbind.data.frame, lapply(datatypes, function(datatype)
    do.call(rbind.data.frame, lapply(derivatives, function(derivative)
      do.call(rbind.data.frame, lapply(crange,function(cycle)
        cbind(datatype = datatype,
              derivative = derivative,
              cycle = cycle,
              details[[datatype]][[derivative]][[paste("c ",cycle)]]$summary
        )
      ))
    ))
  ))


  results_qc <- do.call(rbind.data.frame, lapply(datatypes, function(datatype)
    do.call(rbind.data.frame, lapply(derivatives, function(derivative)
      do.call(rbind.data.frame, lapply(crange,function(cycle)
        cbind(datatype = datatype,
              cycle = cycle,
              derivative = derivative,
              details[[datatype]][[derivative]][[paste("c ",cycle)]]$calib$summary$qc,
              details[[datatype]][[derivative]][[paste("c ",cycle)]]$calib$summary$calib[,-1]
        )
      ))
    ))
  ))

  try(saveRDS(list(results_data = results_data, results_qc = results_qc),
              singleCycleCache), silent = TRUE)
}

c(rows_results_data = nrow(results_data), rows_results_qc = nrow(results_qc))
```

# Panel A — single-cycle estimates from the ΔRn and derRn signals

Panel A is the ΔRn and derRn counterpart of Figure 4B. For each relative cycle
position, the apparent proportion ψₐ and the calibrated proportion ψ_c are
plotted against cycle for all three assays, with the theoretical proportions of
the reference series as horizontal dashed lines. The baseline-subtracted signal
behaves like the raw signal — accuracy improves into the later exponential
cycles — whereas the derivative signal is informative only in a narrow band
around Cᵢ, where the per-cycle fluorescence increase is large enough to carry
the ratio. Each signal is shown twice: the full proportion range, and a zoom on
Δψ = ±0.05 over C₋₅ to C₊₁₀, which is where the differences between the signal
types are actually visible.

The panel is the same plot as Figure 4B — apparent (`aprop`) and calibrated
(`cprop`) proportion per relative cycle, jittered per well and coloured by
theoretical proportion, with the reference proportions as dashed horizontal
lines — evaluated for the other two signals: ΔRn (`deltaRn`, `0D`) and derRn
(`Rn`, `1D`), each as an uncalibrated/calibrated pair. The construction is
taken from the `panel-B` chunk of `ratioPCR_Figure_4.qmd`, not from
`rpcr_plotCalibCycle()`, which the chunk called until 2026-09-14: that function
now draws `delta` (ψ_c − ψ) as a boxplot-plus-jitter deviation view, so it
produced panel-B content under panel A's name. See *Provenance and repairs*.

```{r}
#| label: panel-A
#| echo: true
#| fig-width: 8
#| fig-height: 6

yzoom <- c(-0.05,0.05)
xzoom <- c(-5,10)

qc <- "RSEcprop"
results <- results_data

dfA <- results |>
  tidyr::pivot_longer(cols = c("aprop", "cprop"),
                      names_to = "variable", values_to = "value")

# deltaRn has no first-derivative form; Rn/0D is Figure 4B, not this panel.
dfA <- dfA[!(dfA$datatype == "deltaRn" & dfA$derivative == "1D"), ]
dfA <- dfA[!(dfA$datatype == "Rn" & dfA$derivative == "0D"), ]

dfA$signal <- factor(ifelse(dfA$derivative == "1D", "derRn", "deltaRn"),
                     levels = c("deltaRn", "derRn"))
dfA$variable <- factor(dfA$variable, levels = c("aprop", "cprop"),
                       labels = c("uncalibrated", "calibrated"))

figure_S7_A <- ggplot(data = dfA) +
  geom_hline(yintercept = unique(dfA$tprop), colour = "grey80", linetype = "dashed") +
  geom_jitter(aes(x = cycle, y = value, color = tprop), shape = 20, alpha = 0.7,
              size = 0.2, width = 0.2, height = 0) +
  theme_rpcr() +
  scale_colour_gradientn(colours = rpcr_col_probe) +
  scale_y_continuous(breaks = unique(dfA$tprop)) +
  coord_cartesian(ylim = c(-0.1, 1.1)) +
  facet_grid(target ~ signal + variable) +
  labs(x = expression("cycle relative to C"[i]~"(C"[r]*")"),
       y = expression("proportion ("*psi*")")) +
  theme(axis.text.y = element_text(size = 5),
        legend.position = "bottom")

figure_S7_A
```

The deviation view of the same two signals — the boxplot-plus-jitter rendering
that `rpcr_plotCalibCycle()` returns, full range and zoomed on Δψ = ±0.05 over
C₋₅ to C₊₁₀ — is kept as a diagnostic. It is not panel A and is not written to
`documents/figures`.

```{r}
#| label: panel-A-deviation-diagnostic
#| echo: true
#| fig-width: 8
#| fig-height: 8

p2a <- rpcr_plotCalibCycle(results[results$derivative=="0D"&results$datatype=="deltaRn",],title="deltaRn") + theme_rpcr() + theme(plot.margin = unit(c(0.1,0.2,0.1,0.2), 'lines'))
p2b <- p2a + coord_cartesian(ylim=yzoom,xlim=xzoom)+ ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines'))

p3a <- rpcr_plotCalibCycle(results[results$derivative=="1D"&results$datatype=="Rn",],title="derRn") + theme_rpcr() + theme(plot.margin = unit(c(0.1,0.2,0.1,0.2), 'lines'))
p3b <- p3a + coord_cartesian(ylim=yzoom,xlim=xzoom)+ ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines'))

ggpubr::ggarrange(p2a, p2b, p3a, p3b,
                  ncol = 2, nrow = 2,
                  legend = "bottom",
                  common.legend = TRUE,
                  align = "v")
```

# Panel B — residuals by cycle position and signal type

Panel B is the residual view of the same sweep: the deviation of the calibrated
proportion from the theoretical proportion, Δ = ψ_c − ψ, against relative cycle
position, for the three signal types Rn, ΔRn and derRn and the three assays.
The upper composition covers the full deviation range with a boxplot per cycle
and the individual wells jittered over it, coloured by theoretical proportion —
this is where the cycle positions at which calibration breaks down become
visible. The lower composition is the violin view over the central window,
which shows both the accuracy (offset of the distribution from zero) and the
precision (its width) as a function of cycle selection. The jitter width is
derived from the mean spacing of the cycle grid so that replicates separate
without crossing into the neighbouring cycle.

```{r}
#| label: panel-B-setup
#| echo: true

df <- results_data

# Cycle window of the violin composition. The submitted panel spans C-5 to
# C+10 — its inset brackets start at -5 and +10 on the full-range axis above,
# its lower block carries ticks at -5/0/5/10, and the legend says "C-5 to
# C+10". The window is set here once and used by both the violin chunk and
# the deviation table.
bzoom <- c(-5, 10)


df$signal <- interaction(df$datatype,df$derivative)
df <- df[df$signal!="deltaRn.1D",]


df$signal <- factor(df$signal, levels=c("Rn.0D","deltaRn.0D","Rn.1D"), labels=c("Rn","deltaRn","derRn"))
  jwidth <- unique(df[,"cycle"])
  jwidth <- jwidth[order(jwidth)]
  jwidth <- mean(jwidth[2:(length(jwidth))]- jwidth[1:(length(jwidth)-1)])/5


    plot <- ggplot(data=df) +
        theme_rpcr() +  scale_colour_gradientn(colours = rpcr_col_probe)
```

```{r}
#| label: panel-B-full-range
#| echo: true
#| fig-width: 8
#| fig-height: 7

  plot1 <- plot +
    ggh4x::facet_grid2(target ~ signal, switch = "y", strip = ggh4x::strip_themed(background_y = ggh4x::elem_list_rect(fill = rpcr_col_target)),axes = "x") +
    geom_boxplot(aes(x=cycle,y=delta,group=cycle),outlier.shape = NA) +
    geom_jitter(aes(x=cycle,y=delta, group=interaction(cycle,tprop),col=tprop), # col= if (colplot) as.factor(tprop) else "black",
                             width=jwidth,
                             alpha=0.5,
                             size=1)

plot1
```

```{r}
#| label: panel-B-violins
#| echo: true
#| fig-width: 8
#| fig-height: 7

  plot2 <- ggplot(data=df) +
    theme_rpcr() +  scale_colour_gradientn(colours = rpcr_col_probe)  +
    ggh4x::facet_grid2(target ~ signal,  axes = "all")+
    geom_violin(aes(x=cycle,y=delta,group=cycle)) + #,fill="grey90") +
    geom_jitter(aes(x=cycle,y=delta, group=interaction(cycle,tprop),col=tprop), # col= if (colplot) as.factor(tprop) else "black",
                             width=jwidth,
                             alpha=0.5,
                             size=1)+
    coord_cartesian(ylim=yzoom,xlim=c(bzoom[1]-0.1,bzoom[2]+0.1)) +
    ggtitle("") + theme(plot.margin = unit(c(0.1,0.2,0.5,0.2), 'lines')) +
    scale_x_continuous(breaks=seq(bzoom[1],bzoom[2],by=5))

plot2
```

```{r}
#| label: panel-B-deviation
#| echo: true

# Mean absolute deviation of the calibrated proportion from theoretical, over
# the cycle window shown in the violin composition, per assay and signal type.
knitr::kable(
  df %>%
    filter(cycle >= bzoom[1], cycle <= bzoom[2]) %>%
    group_by(target, signal) %>%
    summarise(n = n(),
              mean_delta = mean(delta, na.rm = TRUE),
              mean_abs_delta = mean(abs(delta), na.rm = TRUE),
              sd_delta = sd(delta, na.rm = TRUE),
              .groups = "drop"),
  digits = 4,
  caption = "Deviation of the calibrated proportion from the theoretical proportion over the cycle window C-5 to C+10 shown in the violin composition, by assay and signal type."
)
```

```{r}
#| label: panel-B-assembled
#| echo: true
#| fig-width: 8
#| fig-height: 12

figure_S7_B <- ggpubr::ggarrange(plot1, plot2, ncol = 1, nrow = 2,
                                 legend = "bottom", common.legend = TRUE)

figure_S7_B
```

# Panel C — RSE heatmaps for the cycle-range slope-fitting method

Panel C is the optimisation surface of the cycle-range method whose design is
Figure 4D. Every combination of window start and window end offset relative to
Cᵢ, from −20 to +20 cycles, is used to fit the slope of the inclusion against
the exclusion signal, the resulting ψₐ is calibrated, and the residual standard
error of the calibrated proportion is recorded per assay. Reading the surface
gives the assay-specific optimal window directly: low-RSE regions are the
windows over which multi-cycle slope fitting is accurate, and the regions where
no value is defined are combinations for which the calibration fit does not
converge.

## Sweep

41 × 41 combinations, of which the 820 with end > start are meaningful windows.
`rtype = "ip"` with `roff = c(start, end)`; `model = "linear"` because a
multi-cycle window is fitted as a straight line through the two-channel signal.

```{r}
#| label: panel-C-sweep
#| echo: true

analysis <- "IP range"

df <- import

datalist <- "Rn"
range1_list <- seq(-20,20,1)
range2_list <- seq(-20,20,1)


if (recalculate || is.null(res[[analysis]]$results_qc)) {

res[[analysis]]$details <-  setNames(lapply(range1_list,function(range1){
               setNames(lapply(range2_list,function(range2) {
                 rpcr_calibrate(df,
                               qual="",
                               output = "details",
                               rtype = "ip", # method to define range
                               roff = c(range1,range2), # offset from range defined by rtype
                               model = "linear" # fitting method: either linear or 2nd degree polynomial (poly)
                   )
                 }), paste0("R_",range2_list))}
               ), paste0("R_",range1_list))



res[[analysis]]$results_data <- do.call(rbind.data.frame, lapply(range1_list, function(range1)
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)
                        cbind(range1 = range1,
                              range2 = range2,
                              res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                              )
                        ))
                      ))


res[[analysis]]$results_qc <-
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)
                        cbind(range1 = range1,
                              range2 = range2,
                               res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                       )
                        ))
                      ))


paraQC <- "RSEcprop"
res[[analysis]]$results_qc2 <-
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)  {
                    settings <- standardSettings
                    settings[,c("rtype","roff")] <- c("ip:ip",paste(c(range1,range2),collapse=":"))
                    temp <-res[[analysis]]$details[[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                    rownames(temp) <- temp$target
                    temp2 <- data.frame(t(temp))[paraQC,]
                    cbind(settings,mean = mean(temp[,paraQC]),temp2)
                    }
               ))
             ))



res[[analysis]]$details <- NULL

}

nrow(res[[analysis]]$results_qc)
```

## Heatmaps

The colour scale is a manual binning of RSE at the **deciles of the RSE
distribution** — `c(0, quantile(value, seq(0, 1, 0.1)))`, eleven bins — computed
over the four maps together and before the `end > start` filter, reversed plasma
so that low RSE is warm; combinations without a defined RSE are drawn grey. Only
windows with end > start are shown, the four maps sit side by side, and the axes
are squared with `coord_fixed()` so that the surface is not distorted.

Data-driven breaks are what the submitted panel used.
The fixed `c(0, 0.01, 0.011, …, 0.019, max)` vector
is a later working
revision: on this sweep it puts almost every window in its first bin, so the
interior gradient and with it the optimum disappear. See *Provenance and
repairs*.

```{r}
#| label: panel-C
#| echo: true
#| fig-width: 8
#| fig-height: 3

analysis <- "IP range"

df <- res[[analysis]]$results_qc


var <- "RSEcprop"

top_mean <- df |>
  dplyr::group_by(range1, range2) |>
  dplyr::summarise(mean = mean(.data[[var]]), .groups = "drop") |>
  dplyr::mutate(mean = mean * 100)

top <- df |>
  dplyr::select(range1, range2, target, value = dplyr::all_of(var)) |>
  tidyr::pivot_wider(
    names_from = target,
    values_from = value
  ) |>
  dplyr::left_join(
    top_mean,
    by = c("range1", "range2")
  )

print(top[which.min(top$mean),])

# Four maps of the panel: the three assays and their mean, built by melting
# results_qc2 over c("mean", the three assays) and facetting on the
# resulting variable; the same long frame is built here from results_qc so
# that the maps and the table share one aggregation. The mean is unscaled
# (the x100 of `top_mean` above is for the printed table only).
dfC <- df[, c("range1", "range2", "target", var)]

dfC_mean <- dfC |>
  dplyr::group_by(range1, range2) |>
  dplyr::summarise(dplyr::across(dplyr::all_of(var), mean), .groups = "drop") |>
  dplyr::mutate(target = "mean")

dfC_all <- rbind(as.data.frame(dfC), as.data.frame(dfC_mean)[, names(dfC)])
dfC_all$target <- factor(dfC_all$target, levels = c("KIF13A", "NUMA1", "SYNE1", "mean"))

# Deciles of the pooled RSE over all four maps, taken before the end > start
# filter is applied. `unique()` only guards against a tied quantile, which
# would make cut() abort; on this sweep the eleven bins are distinct and the
# vector used is unaffected.
breaks_C <- unique(c(0, quantile(dfC_all[[var]], probs = seq(0, 1, by = 0.1),
                                 na.rm = TRUE)))

# The binning is materialised as a column rather than left inside aes(). An
# `aes(fill = cut(get(var), breaks = scale_breaks))` mapping is a quosure: it is
# evaluated when the plot is *drawn*, so the object printed by `psave()` in the
# save-figure chunk picks up whatever `scale_breaks` holds by then — i.e. panel
# D's vector, which is how both saved PDFs came to carry one legend. Failed
# combinations get their own level so the legend reads as in the figure instead
# of "NA".
dfC_all$bin <- cut(dfC_all[[var]], breaks = breaks_C)
levels_C <- levels(dfC_all$bin)
dfC_all$bin <- factor(ifelse(is.na(dfC_all$bin), "calibration failed",
                             as.character(dfC_all$bin)),
                      levels = c(levels_C, "calibration failed"))
col_C <- c(rev(viridis::plasma(n = length(breaks_C)))[seq_along(levels_C)], "grey90")

dfC_all <- dfC_all[dfC_all$range2 > dfC_all$range1, ]

df2 <- dfC_all[!is.na(dfC_all[[var]]), ]
r1 <- c(min(df2$range1, df2$range2), max(df2$range1, df2$range2))

figure_S7_C <- ggplot() +
  geom_tile(dfC_all, mapping = aes(range1, range2, fill = bin), col = "white") +
  scale_fill_manual(values = col_C, drop = FALSE) + theme_rpcr() +
  facet_grid(. ~ target) + coord_fixed() +
  scale_y_continuous(breaks = seq(r1[1], r1[2], by = 10), limits = r1) +
  scale_x_continuous(breaks = seq(r1[1], r1[2], by = 10), limits = r1) +
  labs(x = "relative starting cycle", y = "relative ending cycle") +
  theme(legend.key.size = unit(0.2, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 8),
        strip.background = element_blank(),
        legend.position = "right",
        legend.title = element_blank()) +
  guides(fill = guide_legend(ncol = 1))

figures[[analysis]] <- figure_S7_C

figure_S7_C
```

```{r}
#| label: panel-C-optima
#| echo: true

# Optimal window per assay, and the window minimising the mean across assays.
qcC <- res[["IP range"]]$results_qc
qcC <- qcC[!is.na(qcC[[var]]) & qcC$range2 > qcC$range1, ]

knitr::kable(
  rbind(
    do.call(rbind, lapply(split(qcC, qcC$target), function(d)
      data.frame(scope = unique(d$target), start = d$range1[which.min(d[[var]])],
                 end = d$range2[which.min(d[[var]])], RSEcprop = min(d[[var]])))),
    data.frame(scope = "mean across assays",
               start = top$range1[which.min(top$mean)],
               end = top$range2[which.min(top$mean)],
               RSEcprop = min(top$mean, na.rm = TRUE) / 100)
  ),
  digits = 4, row.names = FALSE,
  caption = "Cycle-range windows (offsets relative to Ci) minimising the residual standard error of the calibrated proportion, per assay and for the mean across assays."
)
```

The legend describes four heatmaps in this panel — the three assays plus the
mean across assays — and all four are now drawn. Here the same long frame is
assembled from `results_qc`, so the map and the optimum in the table above
come from one aggregation.

# Panel D — RSE heatmaps for the derivative-threshold method

Panel D is the optimisation surface of the threshold method whose design is
Figure 4E. The two thresholds — one applied on the rising flank before Cᵢ, one
on the falling flank after it — are swept independently from 0 to 0.99 in 0.05
steps, so that each combination defines a window that adapts to the kinetics of
each individual reaction rather than to a fixed cycle offset. As in panel C the
score is the residual standard error of the calibrated proportion per assay.

## Sweep

```{r}
#| label: panel-D-sweep
#| echo: true

analysis <- "tres double range"

df <- import

datalist <- "Rn"
range1_list <- seq(0,0.99,0.05)
range2_list <- seq(0,0.99,0.05)

if (recalculate || is.null(res[[analysis]]$results_qc)) {

res[[analysis]]$details <-   setNames(lapply(datalist, function(datatype)
               setNames(lapply(range1_list,function(range1){

               setNames(lapply(range2_list,function(range2) {
                 rpcr_calibrate(df,
                               qual="",
                               output = "details",
                               signal = datatype,
                               tres = c(range1,range2),     # treshold value to define dynamic range
                               rtype = "dynamic", # method to define range
                               roff = 0, # offset from range defined by rtype
                               model = "linear" # fitting method: either linear or 2nd degree polynomial (poly)
                   )
                 }), paste0("R_",range2_list))}
               ), paste0("R_",range1_list))
               ), datalist)


res[[analysis]]$results_data <- do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                              res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$summary
                              )
                        ))
                      ))
                    ))

res[[analysis]]$results_qc <-   do.call(rbind.data.frame, lapply(datalist, function(datatype)
                    do.call(rbind.data.frame, lapply(range1_list, function(range1)
                      do.call(rbind.data.frame, lapply(range2_list, function(range2)
                        cbind(datatype = datatype,
                              range1 = range1,
                              range2 = range2,
                               res[[analysis]]$details[[datatype]][[paste0("R_",range1)]][[paste0("R_",range2)]]$calib$summary$qc
                       )
                        ))
                      ))
                    ))

res[[analysis]]$details <- NULL
}

nrow(res[[analysis]]$results_qc)
```

## Heatmaps

Four maps — the three assays and their mean — in one plot on the same binned
scale as panel C: RSE cut at its own deciles, reversed plasma, undefined
combinations grey. Threshold 1 is on the x-axis and threshold 2 on the y-axis,
both as percentages (the sweep stores them as fractions 0–0.95, and the panel
multiplies by 100), and `coord_fixed()` squares the maps.

```{r}
#| label: panel-D
#| echo: true
#| fig-width: 8
#| fig-height: 3

analysis <- "tres double range"

df <- res[[analysis]]$results_qc

var <- "RSEcprop"

top_mean <- df |>
  dplyr::group_by(datatype, range1, range2) |>
  dplyr::summarise(mean = mean(.data[[var]]), .groups = "drop") |>
  dplyr::mutate(mean = mean * 100)

top <- df |>
  dplyr::select(datatype, range1, range2, target, value = dplyr::all_of(var)) |>
  tidyr::pivot_wider(
    names_from = target,
    values_from = value
  ) |>
  dplyr::left_join(
    top_mean,
    by = c("datatype", "range1", "range2")
  )


print(top[which.min(top$mean),])


# Four maps in one plot: the three assays and their mean, melted and
# facetted into a single grid. The mean is the unscaled counterpart of
# `top_mean` above, so the map and the optimum in the table below agree.
dfD <- df[, c("range1", "range2", "target", var)]

dfD_mean <- dfD |>
  dplyr::group_by(range1, range2) |>
  dplyr::summarise(dplyr::across(dplyr::all_of(var), mean), .groups = "drop") |>
  dplyr::mutate(target = "mean")

dfD_all <- rbind(as.data.frame(dfD), as.data.frame(dfD_mean)[, names(dfD)])
dfD_all$target <- factor(dfD_all$target, levels = c("KIF13A", "NUMA1", "SYNE1", "mean"))

breaks_D <- unique(c(0, quantile(dfD_all[[var]], probs = seq(0, 1, by = 0.1),
                                 na.rm = TRUE)))

dfD_all$bin <- cut(dfD_all[[var]], breaks = breaks_D)
levels_D <- levels(dfD_all$bin)
dfD_all$bin <- factor(ifelse(is.na(dfD_all$bin), "calibration failed",
                             as.character(dfD_all$bin)),
                      levels = c(levels_D, "calibration failed"))
col_D <- c(rev(viridis::plasma(n = length(breaks_D)))[seq_along(levels_D)], "grey90")

figure_S7_D <- ggplot() +
  geom_tile(dfD_all, mapping = aes(range1 * 100, range2 * 100, fill = bin),
            col = "white") +
  scale_fill_manual(values = col_D, drop = FALSE) + theme_rpcr() +
  facet_grid(. ~ target) + coord_fixed() +
  labs(x = "treshold 1 [%]", y = "treshold 2 [%]") +
  theme(legend.key.size = unit(0.2, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 8),
        strip.background = element_blank(),
        legend.position = "right",
        legend.title = element_blank()) +
  guides(fill = guide_legend(ncol = 1))

figures[[analysis]] <- figure_S7_D

figure_S7_D
```

```{r}
#| label: panel-D-optima
#| echo: true

qcD <- res[["tres double range"]]$results_qc
qcD <- qcD[!is.na(qcD[[var]]), ]

knitr::kable(
  rbind(
    do.call(rbind, lapply(split(qcD, qcD$target), function(d)
      data.frame(scope = unique(d$target), threshold1 = d$range1[which.min(d[[var]])],
                 threshold2 = d$range2[which.min(d[[var]])], RSEcprop = min(d[[var]])))),
    data.frame(scope = "mean across assays",
               threshold1 = top$range1[which.min(top$mean)],
               threshold2 = top$range2[which.min(top$mean)],
               RSEcprop = min(top$mean, na.rm = TRUE) / 100)
  ),
  digits = 4, row.names = FALSE,
  caption = "Derivative thresholds minimising the residual standard error of the calibrated proportion, per assay and for the mean across assays. threshold1 is applied before Ci, threshold2 after Ci."
)
```

As in panel C, the mean-across-assays map is the fourth map of the panel, and
all four share one scale, so the maps are read against one another: the assay
differences the legend text describes ("warm colours indicate optimal threshold
combinations") are differences between maps, which a per-map scale would hide.
The earlier per-map `scale_fill_viridis()` version of this chunk — four
independently scaled maps assembled with `ggpubr::ggarrange()` — is superseded;
it inverted the colour direction as well, since viridis puts its bright end at
the *high* RSE of the failing corner.

## Appendix — the "Fixed cycle range" sweep

Chunk `Figure_005` renders a second family of
heatmaps from the `res[["Fixed cycle range"]]` sweep, using a rescaled viridis
scale that saturates above RSE = 0.02. That sweep is **not** the
threshold sweep of panel D: chunk `range_005` computes it with `rtype = "ip"`
and `roff = c(range1, range2)` over the same −20…+20 cycle grid as panel C, for
two signal types and two derivative settings — 2 × 2 × 41 × 41 = 6 724
calibrations. It is therefore a signal-type-resolved variant of panel C's
surface rather than a threshold surface.

Because the panel map cites it alongside `range10` for panel D, it is rendered
here when the cache provides it, clearly marked as an alternative surface. It is
never recomputed by this document — the sweep is an order of magnitude more
expensive than any other in the figure set, and nothing in the S7 legend
requires it.

```{r}
#| label: appendix-fixed-cycle-range
#| echo: true
#| fig-width: 9
#| fig-height: 3.5

analysis <- "Fixed cycle range"

if (is.null(res[[analysis]]$results_qc)) {

  cat("Not rendered: res[['Fixed cycle range']] is not in the cache and this\n",
      "document does not recompute it (2 x 2 x 41 x 41 = 6724 calibrations).\n",
      "To populate it, run chunk `range_005` with\n",
      "recalculate = TRUE and re-save the results cache.\n", sep = "")

} else {

  df <- na.omit(res[[analysis]]$results_qc)

  var <- "RSEcprop"

  fixed_range_maps <- lapply(genelist, function(target){
    scaly <- 0.02
    ggplot(df[df$target==target,], aes(range1, range2, fill= get(var))) +
      geom_tile() +
      scale_fill_viridis(rescaler = function(x, to = c(0, 1), from = NULL) {
        ifelse(x<scaly,
               scales::rescale(x, to = to,
                               from = c(min(x, na.rm = TRUE), scaly)),
               1)})
  })

  figures[[analysis]] <- fixed_range_maps
  print(ggpubr::ggarrange(plotlist = fixed_range_maps, ncol = 3, nrow = 1,
                          labels = genelist, legend = "bottom",
                          font.label = list(size = 9, face = "bold", color = "black")))
}
```

# Colour-scale audit

The two break vectors are the deciles of the two sweeps, so they are a direct
check on the data behind the panels: the legend of the submitted figure prints
the same eleven intervals. Both vectors, and the published ones read off the
submitted legend, are written to `figures/components/S7_scale_audit.txt`.

```{r}
#| label: scale-audit
#| echo: true

published_C <- c(0, 0.00417, 0.00505, 0.0056, 0.00585, 0.00637, 0.00715,
                 0.00763, 0.00931, 0.0141, 0.0273, 0.287)
published_D <- c(0, 0.00411, 0.00558, 0.00604, 0.0064, 0.00669, 0.00702,
                 0.00736, 0.00767, 0.00812, 0.00863, 0.0128)

audit <- data.frame(
  panel  = c("C", "C", "D", "D"),
  source = c("this render", "submitted", "this render", "submitted"),
  n_breaks = c(length(breaks_C), length(published_C),
               length(breaks_D), length(published_D)),
  breaks = c(paste(signif(breaks_C, 3), collapse = ", "),
             paste(signif(published_C, 3), collapse = ", "),
             paste(signif(breaks_D, 3), collapse = ", "),
             paste(signif(published_D, 3), collapse = ", "))
)

counts <- data.frame(
  panel = c("C", "D"),
  n_values = c(nrow(dfC_all), nrow(dfD_all)),
  n_failed = c(sum(dfC_all$bin == "calibration failed"),
               sum(dfD_all$bin == "calibration failed")),
  max_RSE = c(max(dfC_all[[var]], na.rm = TRUE), max(dfD_all[[var]], na.rm = TRUE))
)

writeLines(c(capture.output(print(audit, row.names = FALSE)), "",
             capture.output(print(counts, row.names = FALSE))),
           file.path(cpath, "S7_scale_audit.txt"))

knitr::kable(audit, caption = "Colour-scale breaks of panels C and D as rendered here, against the intervals printed in the legend of the submitted figure.")
counts
```

# Save figure

```{r}
#| label: save-figure
#| echo: true

psave(filename = "ratioPCR_figure_S7A", plot = figure_S7_A, path = fpath,
      device = "pdf", width = 20, height = 14, units = "cm", dpi = 300)
psave(filename = "ratioPCR_figure_S7B", plot = figure_S7_B, path = fpath,
      device = "pdf", width = 20, height = 30, units = "cm", dpi = 300)
psave(filename = "ratioPCR_figure_S7C", plot = figure_S7_C, path = fpath,
      device = "pdf", width = 20, height = 8, units = "cm", dpi = 300)
psave(filename = "ratioPCR_figure_S7D", plot = figure_S7_D, path = fpath,
      device = "pdf", width = 20, height = 8, units = "cm", dpi = 300)

list.files(fpath, pattern = "^ratioPCR_figure_S7", full.names = FALSE)
```

# Provenance and repairs

**The fourth map of panels C and D** — the mean across the three assays, which
both panels' legends describe and which this document previously recorded as
having no plotting code — comes from the sweeps that were part of
Figure 5: the cycle-range sweep (panel C) and the threshold sweep (panel D).
Both build a long frame over `c("mean", "KIF13A", "NUMA1", "SYNE1")` from a
`results_qc2` table and facet on that variable, so the mean map is a facet of
the same plot as the assay maps.
The long frame is rebuilt from `results_qc`; `results_qc2` is carried in the
cached results for the cycle-range sweep but not for the threshold sweep, and
the two routes give the same 6724 and 1600 values respectively. The aggregation
is identical to the `top_mean` used for the optimum tables, without the ×100
that those tables apply for display.

The colour treatment is **no longer** each panel's own (superseded 2026-09-14).
The first transplant kept the fixed RSE binning in C and the per-map
`scale_fill_viridis()` in D and recorded the quantile scaling as an
unrelated difference. It is not unrelated: the
scale is what the panel shows. Both panels now use the archived scripts'
construction.

**Panel A** is now built directly. Repairs:

0.  **`rpcr_plotCalibCycle()` no longer draws this panel** (2026-09-14).
    `R/rpcr_plotCalibCycle.R` maps `y = delta` — the deviation ψ_c − ψ — as
    `geom_boxplot()` plus jitter, faceted by target. The submitted panel A is a
    *proportion* view: `aprop` and `cprop` per relative cycle, coloured by
    theoretical proportion, with the eleven reference proportions as dashed
    lines and the y-axis broken at those proportions. Calling the function
    therefore produced panel-B content under panel A's name — two panels of the
    same figure showing the same quantity — with `get(var)` as the x-axis title
    and `deltaPSI` as the y-axis title. The panel is now assembled from the
    `panel-B` chunk of `ratioPCR_Figure_4.qmd`, which is the same plot for the
    Rn signal and is the version that went into the Figure 4 panel file,
    evaluated here for ΔRn and derRn. Whether `rpcr_plotCalibCycle()` should be
    restored to its pre-rename behaviour is a package decision and is not made
    here; the deviation view it returns is kept as a diagnostic under panel A.

1.  **Panel composition.** The chunk arranges all six plots (Rn, ΔRn, derRn ×
    full/zoom) into `p4a` and then into a nine-row `p4` scaffold of
    `nullplot`s, extracted legends and a page title ("Calculation with single
    cycle relative to inflection point") — a full-page exploratory layout, not
    a figure panel, and it labels six plots with four labels
    (`labels = c("A","B","C","D")`). Supplementary Figure S7A is the ΔRn and
    derRn pair only, per its legend, with the Rn pair being Figure 4B; the panel
    is one plot faceted `target ~ signal + variable`, i.e. three assay rows by
    ΔRn/derRn × uncalibrated/calibrated, as in the submitted figure. The
    `nullplot` scaffold, the page title and the `p1a`/`p1b` Rn pair are dropped.

2.  **Deprecated legend extraction dropped.** `p_legend_a`/`p_legend_b1`/
    `p_legend_b2` use `ggpubr::get_legend()` together with
    `guides(shape = FALSE)` and `guides(color = FALSE)`; the logical form of
    `guides()` has been deprecated since ggplot2 3.3.4 and errors on current
    versions. They existed only to feed the dropped scaffold, so they are
    removed rather than rewritten.

3.  **Dead code removed.** The chunk filters `results` and then immediately
    reassigns `results <- results_data`, discarding the filter; the filtered
    copies `res_qc1`, `failed` and the destructive overwrite of `results_qc`
    (which would corrupt the object panels B–D read) are dropped. `p5a`/`p5b`,
    the log-scale RSE plot and the failed-calibration strip, are the
    exploratory precursors of Figure 4C and are not part of S7.

**Panel B**'s `signal`
factor construction, the `jwidth` derivation, `plot`, `plot1` and `plot2` are
unchanged, including the `ggh4x` strip theming. Repairs: `yzoom` is defined in
this document, and
`plot1b` — a duplicate of `plot1` with blanked strips, never used — is dropped.

**Violin window corrected** (2026-09-14). This document previously kept
`coord_cartesian(xlim = c(-1.1, 7.1))` with `scale_x_continuous(breaks = -1:7)`
and recorded the legend's "C₋₅ to C₊₁₀" as the
discrepancy, on the assumption that the code was what produced the submitted
panel. The submitted panel settles it the other way: the inset brackets leave
the full-range axis at −5 and +10, the lower block is ticked at −5/0/5/10, and
it holds sixteen violins per facet. The window is now `bzoom <- c(-5, 10)`, set
once and used by both the violin composition and the deviation table, so the
table still covers exactly what the panel shows.

**Panel C** is transplanted from chunk `range11` (line 1063) and the tile chunk
at line 1134: the sweep, `results_qc2` with its `standardSettings` encoding, the
`range2 > range1` filter, the axis squaring and `coord_fixed()` are unchanged.
Repairs:

0.  **Colour scale and facet direction taken from the script that produced the
    submitted panel** (2026-09-14). The tile chunk bins RSE on a fixed
    `c(0, 0.01, 0.011, …, 0.019, max)` vector and facets `target ~ .`; the
    submitted panel bins at `c(0, quantile(value, seq(0, 1, 0.1)))` — eleven
    bins over the pooled four maps — and facets `. ~ variable`. The fixed
    vector is not a stylistic variant: on this sweep it collapses the whole
    low-RSE region into its first bin, so the optimum the panel exists to show
    is not visible, and the legend printed 9 intervals in the 0.010–0.019
    decade where the submitted figure prints 11 spanning 0.00417–0.287. The
    document now prints its break vector so that it can be compared against
    the submitted legend directly.

1.  **Recompute guard widened** from `if (recalculate)` to
    `if (recalculate || is.null(res[[analysis]]$results_qc))`, so the panel
    builds even when the cache does not carry this analysis.

2.  **Dead code that only worked by accident removed.** The tile chunk opens
    with a `top` assignment that loops over `datalist` and subsets on
    `df$datatype` — a column the `range11` sweep never creates. It evaluates to
    a zero-row frame and is then overwritten by the `pivot_wider()` version two
    statements later. Only the surviving definition is kept. `datalist` is set
    to `"Rn"` in the sweep chunk as in the notebook, rather than inherited from
    `Figure_settings.R` (where it is `c("F","bF")`).

3.  **Bare output path.** `ggsave(paste0(fpath, analysis, ".png"), ...)`
    concatenates without a separator; replaced by `psave("ratioPCR_figure_S7C", ...)`.

4.  **Additive readout.** The optimal window per assay and for the mean across
    assays is tabulated so the legend's "optimal window of cycle positions" is
    a printed value rather than a claim.

**Panel D**: the threshold sweep is transplanted from chunk `range10`
(line 945) unchanged. Repairs: the recompute guard is widened as for panel C;
the unused `results_qc_na <- results_qc %>% replace(is.na(.), 1)` line — which
reaches across into the single-cycle `results_qc` and is never read — is
dropped, as is the accidental `top` assignment described above.

The **tile maps** deliberately depart from the earlier continuous-scale
rendering (2026-09-14), in which one `scale_fill_viridis()` map per assay was
assembled from four separate plots with `ggpubr::ggarrange()`. That is not the
submitted panel, and the mismatch is not only cosmetic: four independent scales
cannot be compared across assays, the bright end of viridis marks the *highest*
RSE (so the failing threshold corner reads as the optimum), the threshold axes
stay on the sweep's 0–0.95 fraction instead of the panel's 0–100 %, and the
assay order follows `genelist` (NUMA1, SYNE1, KIF13A) rather than the figure's
KIF13A, NUMA1, SYNE1. The panel is built here as one plot, four facets, RSE cut
at its deciles, reversed plasma, `range * 100` on both axes — which reproduces
the submitted legend (eleven bins from 0.00411 to 0.0128) and its grey
"calibration failed" corner.

**Documented gaps in this figure.** The mean-across-assays heatmap named in the
legend of both panel C and panel D is drawn since 2026-09-14, from the archived
sweep scripts (first section of *Provenance and repairs*); it is no longer a
gap. The `Figure_005` /
"Fixed cycle range" surface is rendered only from cache and is, on inspection
of chunk `range_005`, a cycle-range sweep resolved by signal type rather than
the threshold surface of panel D.

# Manuscript figure

The files this document writes into `documents/figures`, re-printed here at
the size they are written at, so that the rendered document shows the panels
as they appear in the manuscript rather than only the analysis-sized views
above. The final chunk checks that every file is on disk.

**`ratioPCR_figure_S7A.pdf`** — panel A, 20 × 14 cm.

```{r}
#| label: manuscript-figure-S7A
#| echo: false
#| fig-width: 7.87
#| fig-height: 5.51

figure_S7_A
```

**`ratioPCR_figure_S7B.pdf`** — panel B, 20 × 30 cm.

```{r}
#| label: manuscript-figure-S7B
#| echo: false
#| fig-width: 7.87
#| fig-height: 11.81

figure_S7_B
```

**`ratioPCR_figure_S7C.pdf`** — panel C, 20 × 8 cm.

```{r}
#| label: manuscript-figure-S7C
#| echo: false
#| fig-width: 7.87
#| fig-height: 3.15

figure_S7_C
```

**`ratioPCR_figure_S7D.pdf`** — panel D, 20 × 8 cm.

```{r}
#| label: manuscript-figure-S7D
#| echo: false
#| fig-width: 7.87
#| fig-height: 3.15

figure_S7_D
```

```{r}
#| label: manuscript-figure-files
#| echo: false

written <- c("ratioPCR_figure_S7A.pdf", "ratioPCR_figure_S7B.pdf", "ratioPCR_figure_S7C.pdf", "ratioPCR_figure_S7D.pdf")
data.frame(file = written,
           exists = file.exists(file.path(fpath, written)),
           row.names = NULL)
```

# Reproducibility

```{r}
#| label: session-info
#| echo: true

if (requireNamespace("sessioninfo", quietly = TRUE)) {
  sessioninfo::session_info()
} else {
  print(sessionInfo())
}

if (requireNamespace("renv", quietly = TRUE)) {
  try(print(renv::status()), silent = TRUE)
}
```
