---
title: "ratioPCR Figure S9 — Calibration quality improves with increasing number of reference standards and stabilises above seven"
format:
  html:
    embed-resources: true
    code-background: true
    code-fold: true
    code-tools: true
    toc: true
    toc-depth: 3
    number-sections: true
editor: visual
execute:
  warning: false
  message: false
---

<!-- ============================================================= -->
<!-- ratioPCR_Figure_S9.qmd                                         -->
<!-- Self-contained rebuild of Supplementary Figure S9              -->
<!-- (ISCIENCE-D-26-02393_R2).                                      -->
<!-- The combination sweep is expensive; it is cached in            -->
<!-- documents/extdata/RefNum_calculation.rds and only recomputed   -->
<!-- when RPCR_RECALCULATE=TRUE.                                    -->
<!-- ============================================================= -->

# Figure legend

> **Supplementary Figure S9: Calibration quality improves with increasing
> number of reference standards and stabilises above seven**
>
> Violin plots showing the distribution of root mean square error (RMSE) of
> calibrated inclusion proportions across all possible combinations of 4 to 11
> reference standards drawn from an evenly spaced series (0% to 100% inclusion
> in 10% steps) using the KIF13A assay. For each combination, calibration
> parameters were estimated from one reference series and RMSE was evaluated on
> a held-out cross-validation series. The width of each violin reflects the
> density of RMSE values across all tested combinations at that standard
> number. With four reference standards — the theoretical minimum required to
> fit the four-parameter calibration model — RMSE is highly variable (range
> ~0.01–0.43), indicating that calibration quality depends strongly on which
> four standards happen to be selected. Variability decreases consistently with
> increasing standard number, with worst-case RMSE falling below 0.05 at nine
> or more standards. From nine standards onwards, the RMSE distribution
> collapses to a narrow band around ~0.03, indicating robust calibration
> independent of specific standard selection. At eleven standards only one
> combination exists, appearing as a single point. These results support a
> practical recommendation of a minimum of five reference standards distributed
> across the full ratio range, with seven or more preferred for highest
> robustness.

------------------------------------------------------------------------

# Setup

The four-parameter calibration model needs at least four standards, but which
four are chosen matters. The sweep behind this figure fits the model on every
subset of size 4 to 11 drawn from the eleven-point KIF13A reference series and
evaluates the resulting calibration on a second, independently prepared series
of the same mixtures — a held-out evaluation, so that the reported RMSE
measures transferability of the calibration rather than its fit to its own
standards. Two reference series are available on the plate (`laKIF13A`, series
A, and `lbKIF13A`, series B); each is used in turn as the fitting series and
each combination is evaluated both on itself and on the other, giving the four
classes `AA`, `AB`, `BA`, `BB` (first letter = fitted on, second = evaluated
on).

The sweep is expensive — the full enumeration is
`sum(choose(11, 4:11)) = 1816` combinations per series, each requiring a model
fit and a batch refit. It is therefore cached in
`documents/extdata/RefNum_calculation.rds` and read from there by default. Set
the environment variable `RPCR_RECALCULATE=TRUE` before rendering to run the
sweep from the raw plate data instead.

```{r}
#| label: setup

pp <- here::here()
source(file.path(pp, "documents", "scripts", "Figure_settings.R"))

# Output naming (added when the figure documents were put on one convention).
# documents/figures holds the manuscript assets, one file per panel or panel
# range, named ratioPCR_figure_<figure><panel>. documents/figures/components
# holds the pieces that were assembled into a panel outside R, the diagnostic
# plots and the alternative views, under their working names.
cpath <- file.path(fpath, "components")
dir.create(fpath, recursive = TRUE, showWarnings = FALSE)
dir.create(cpath, recursive = TRUE, showWarnings = FALSE)

ppath <- file.path(pp, "documents", "extdata", "dPCR")

# Cache location. The source script pointed at documents/data, where the file
# no longer lives; it has been restored to documents/extdata.
refnum_cache <- file.path(pp, "documents", "extdata", "RefNum_calculation.rds")

# Recompute switch. Default FALSE: read the cached sweep.
recalculate <- as.logical(Sys.getenv("RPCR_RECALCULATE", "FALSE"))

target <- "KIF13A"
```

------------------------------------------------------------------------

# Combination sweep

The block below is the sweep itself. It runs only when `recalculate` is `TRUE`;
otherwise the cached result is read.

```{r}
#| label: refnum-sweep

if (recalculate) {

# load data
import <- rpcr_wrap(rpcr_readTF(file.path(ppath, "rPCR_20250916_JK KIF13A dPCR comparison_20250918_125346.xlsx")))
import0 <- import[import$target %in% c("laKIF13A", "lbKIF13A"), ]
import0 <- import0[import0$sample != "NTC", ]

# Calibration quality metrics for theoretical vs calibrated proportions
# df must contain:
#   tprop = theoretical proportion
#   cprop = calibrated proportion

calc_calibration_metrics <- function(df, eps = 1e-4) {
  stopifnot(all(c("tprop", "cprop") %in% names(df)))

  d <- df[is.finite(df$tprop) & is.finite(df$cprop), ]

  # Bound check
  if (any(d$tprop < 0 | d$tprop > 1 | d$cprop < 0 | d$cprop > 1)) {
    warning("Some proportions are outside [0,1]. Please check input data.")
  }

  # Proportion-scale residuals
  d$error <- d$cprop - d$tprop
  d$abs_error <- abs(d$error)
  d$sq_error <- d$error^2

  # Metrics on proportion scale
  prop_metrics <- data.frame(
    n = nrow(d),
    bias = mean(d$error),
    MAE = mean(d$abs_error),
    RMSE = sqrt(mean(d$sq_error)),
    max_abs_error = max(d$abs_error)
  )

  # Logit-scale metrics, excluding theoretical or calibrated boundary values
  d_logit <- subset(d, tprop > 0 & tprop < 1 & cprop > 0 & cprop < 1)

  logit <- function(x) log(x / (1 - x))
  clamp <- function(x) pmin(pmax(x, eps), 1 - eps)

  if (nrow(d_logit) > 0) {
    d_logit$tlogit <- logit(d_logit$tprop)
    d_logit$clogit <- logit(d_logit$cprop)
    d_logit$logit_error <- d_logit$clogit - d_logit$tlogit

    logit_metrics <- data.frame(
      log_n = nrow(d_logit),
      log_bias = mean(d_logit$logit_error),
      log_MAE = mean(abs(d_logit$logit_error)),
      log_RMSE = sqrt(mean(d_logit$logit_error^2)),
      log_max_abs_error = max(abs(d_logit$logit_error))
    )
  } else {
    logit_metrics <- data.frame(
      log_n = 0,
      log_bias = NA_real_,
      log_MAE = NA_real_,
      log_RMSE = NA_real_,
      log_max_abs_error = NA_real_
    )
  }
  tlogit <- logit(clamp(d$tprop))
  clogit <- logit(clamp(d$cprop))
  err <- clogit - tlogit

  logit_clamp_metrics <- data.frame(
    logc_n = length(err),
    logc_bias = mean(err),
    logc_MAE = mean(abs(err)),
    logc_RMSE = sqrt(mean(err^2)),
    logc_max_abs_error = max(abs(err))
  )
  cbind(prop_metrics, logit_metrics, logit_clamp_metrics)
}

res <- list()

#### Series A as calibration samples

my_numbers <- na.omit(unique(rpcr_calibDetect(import0[import0$target == "laKIF13A", ])$tprop))

all_combinations <- lapply(4:length(my_numbers), function(m) {
  # Generate combinations of size m, returning a list of vectors
  combn(my_numbers, m, simplify = FALSE)
})

final_list <- unlist(all_combinations, recursive = FALSE)

res$A <- do.call(rbind, lapply(seq_along(final_list), function(m) {
  calib <- unlist(rpcr_calibrate(import0[import0$target == "laKIF13A" & import0$sample %in% paste0("Cal", final_list[[m]] * 100), ])[, -1])
  if (!anyNA(calib[c("a", "ri", "re", "g")])) {
  qc <- rpcr_calibDetect(rpcr_batchFit(import0[import0$target == "laKIF13A", ], calib = calib))
  A <- cbind(ID = m, class = "AA", calc_calibration_metrics(qc), res_num = length(final_list[[m]]), refs = paste(final_list[[m]], collapse = ", "))
  qc <- rpcr_calibDetect(rpcr_batchFit(import0[import0$target == "lbKIF13A" & import0$well != "G1", ], calib = calib))
  B <- cbind(ID = m, class = "AB", calc_calibration_metrics(qc), res_num = length(final_list[[m]]), refs = paste(final_list[[m]], collapse = ", "))
  rbind(A, B)
  }}))

#### Series B as calibration samples

my_numbers <- na.omit(unique(rpcr_calibDetect(import0[import0$target == "lbKIF13A", ])$tprop))

all_combinations <- lapply(4:length(my_numbers), function(m) {
  # Generate combinations of size m, returning a list of vectors
  combn(my_numbers, m, simplify = FALSE)
})

final_list <- unlist(all_combinations, recursive = FALSE)

res$B <- do.call(rbind, lapply(seq_along(final_list), function(m) {
  calib <- unlist(rpcr_calibrate(import0[import0$target == "lbKIF13A" & import0$sample %in% paste0("Cal", final_list[[m]] * 100), ])[, -1])
  if (!anyNA(calib[c("a", "ri", "re", "g")])) {
    qc <- rpcr_calibDetect(rpcr_batchFit(import0[import0$target == "laKIF13A", ], calib = calib))
    A <- cbind(ID = m, class = "BA", calc_calibration_metrics(qc), res_num = length(final_list[[m]]), refs = paste(final_list[[m]], collapse = ", "))
    qc <- rpcr_calibDetect(rpcr_batchFit(import0[import0$target == "lbKIF13A", ], calib = calib))
    B <- cbind(ID = m, class = "BB", calc_calibration_metrics(qc), res_num = length(final_list[[m]]), refs = paste(final_list[[m]], collapse = ", "))
    rbind(A, B)
  }}
))

saveRDS(res, refnum_cache)
}

if (!recalculate) res <- readRDS(refnum_cache)

cat("Sweep source:", if (recalculate) "recomputed from raw plate data" else refnum_cache, "\n")
cat("Rows per fitting series: A =", nrow(res$A), ", B =", nrow(res$B), "\n")
cat("Classes present:", paste(sort(unique(c(res$A$class, res$B$class))), collapse = ", "), "\n")
```

------------------------------------------------------------------------

# Panel — RMSE against number of reference standards

Each violin is the distribution of RMSE over all combinations of a given size.
The relevant comparison is not the mean but the upper tail: a calibration
strategy is only safe if the *worst* combination of that size is still
acceptable, because in practice the standards are chosen once and their
identity is not known to be favourable. The number of combinations is
`choose(11, k)`, so the violins are widest in the middle of the range and
collapse to a single point at eleven standards, where only one combination
exists.

The class breakdown below is built first, because it separates the two
evaluations that the figure pools: `AA`/`BB` evaluate a calibration on the same
series it was fitted on, `AB`/`BA` evaluate it on the held-out series. The
legend describes the held-out case.

```{r}
#| label: figS9-class-breakdown
#| fig-width: 5
#| fig-height: 6

res0 <- rbind(res$A, res$B)

pRMSE <- ggplot(data = res0) + geom_violin(aes(x = res_num, y = RMSE, group = res_num)) + facet_grid(class ~ .) + theme_rpcr()

pRMSE
```

The submitted panel is built from the series-A sweep, expanded by reference
value. `aha` re-emits each combination once for every standard it contains,
which is how the source script prepared the data for a per-standard
breakdown; the expansion leaves the range of RMSE at each combination size
unchanged but weights the violin density by combination size. The expansion is
kept as written, because this is the object that produced the submitted figure.

```{r}
#| label: figS9-panel
#| fig-width: 4.5
#| fig-height: 2.8

refs <- c(0, 0.05, 0.15, 0.25, 0.35, 0.5, 0.65, 0.75, 0.85, 0.95, 1)

res0 <- res$A

aha <- do.call(rbind, lapply(unique(res0$res_num), function(num) {
  df <- res0[res0$res_num == num, ]
  do.call(rbind, lapply(refs, function(x) cbind(prop = x, df[grepl(paste(x), df$refs, fixed = TRUE), ])))
}))

final <- ggplot(data = aha) + geom_violin(aes(x = res_num, y = RMSE, group = res_num)) +
  scale_x_continuous("number of reference samples", labels = as.character(unique(aha$res_num)), breaks = unique(aha$res_num)) +
  theme_rpcr()

final
```

------------------------------------------------------------------------

# Numbers quoted in the legend

The statements in the legend are read off the sweep here rather than restated.
They are computed on the unexpanded series-A result, so that each combination
counts once.

```{r}
#| label: figS9-numbers

sweep_summary <- res$A %>%
  group_by(res_num, class) %>%
  summarise(n_combinations = dplyr::n(),
            min_RMSE    = min(RMSE, na.rm = TRUE),
            median_RMSE = median(RMSE, na.rm = TRUE),
            max_RMSE    = max(RMSE, na.rm = TRUE),
            .groups = "drop")

knitr::kable(sweep_summary, digits = 4,
             caption = "RMSE by number of reference standards, series A fit (AA = self-evaluated, AB = held-out series)")

held_out <- sweep_summary[sweep_summary$class == "AB", ]

cat("Four standards, held-out evaluation: RMSE range ",
    sprintf("%.3f", held_out$min_RMSE[held_out$res_num == 4]), " to ",
    sprintf("%.3f", held_out$max_RMSE[held_out$res_num == 4]),
    " over ", held_out$n_combinations[held_out$res_num == 4], " combinations.\n", sep = "")

below <- held_out$res_num[held_out$max_RMSE < 0.05]
cat("Standard numbers at which the worst-case held-out RMSE is below 0.05: ",
    if (length(below)) paste(below, collapse = ", ") else "none", "\n", sep = "")

nine_plus <- held_out[held_out$res_num >= 9, ]
cat("Nine standards and above, held-out median RMSE: ",
    paste(sprintf("%d standards: %.3f", nine_plus$res_num, nine_plus$median_RMSE),
          collapse = "; "), "\n", sep = "")

cat("Combinations at eleven standards: ",
    held_out$n_combinations[held_out$res_num == 11], "\n", sep = "")
```

------------------------------------------------------------------------

# Save figure

The source script wrote into `documents/figures/dPCR`, a subdirectory that is
not created anywhere in the repository; the write is routed here through
`psave()` into `fpath` (`documents/figures`), as for every other figure
document. The file name is otherwise unchanged.

```{r}
#| label: save-figures

dir.create(fpath, recursive = TRUE, showWarnings = FALSE)

psave("ratioPCR_figure_S9", final, width = 11.4, height = 7, dpi = 600)

cat("Figure S9 output written to:", fpath, "\n")
```

------------------------------------------------------------------------

# Provenance and repairs

| Element | Source | Repairs applied |
|----|----|----|
| Panel | object `final` → `ratioPCR_figure_S9.pdf` | Cache path corrected from `documents/data/RefNum_calculation.rds` to `documents/extdata/RefNum_calculation.rds` (both read and write). The hardcoded `recalculate <- FALSE` replaced by the `RPCR_RECALCULATE` environment switch, default `FALSE`. Output directory changed from the non-existent `documents/figures/dPCR` to `fpath`. `print()` progress calls inside the sweep loops removed. Interactive dead code removed: the commented-out "all series" sweep, seven unassigned exploratory `ggplot()` calls, the `resx[order(...)][1:10,]` inspection blocks, and the `test`/`psym` symmetry analysis, none of which contributes to the figure. Plotting code for `pRMSE` and `final` unchanged. |

Two points that the reader of this figure should know, neither of which was
changed:

1. **The submitted panel pools `AA` and `AB`.** `aha` is built from `res$A`,
   which contains both the self-evaluated (`AA`) and the held-out (`AB`)
   rows, and `final` does not facet by class. The legend describes the
   held-out evaluation only. The class breakdown plotted above shows the two
   separately; the pooled panel is retained because it is the figure that was
   submitted.
2. **`aha` duplicates rows.** Each combination is re-emitted once per reference
   value it contains, so a combination of eleven standards contributes about
   eleven rows and a combination of four about four. The match is a substring
   test on the comma-joined label (`grepl(..., fixed = TRUE)`), so the
   duplication factor is not exactly the combination size — `0` matches every
   label containing a zero and `1` also matches `0.15`. Minimum, maximum and
   range at each combination size are unaffected either way; only the violin
   *density* is reweighted. The numbers section above is therefore computed on
   the unexpanded `res$A`.

------------------------------------------------------------------------

# Manuscript figure

The files this document writes into `documents/figures`, re-printed here at
the size they are written at, so that the rendered document shows the panels
as they appear in the manuscript rather than only the analysis-sized views
above. The final chunk checks that every file is on disk.

**`ratioPCR_figure_S9.pdf`** — the whole figure, 11.4 × 7 cm.

```{r}
#| label: manuscript-figure-S9
#| echo: false
#| fig-width: 4.49
#| fig-height: 2.76

final
```

```{r}
#| label: manuscript-figure-files
#| echo: false

written <- c("ratioPCR_figure_S9.pdf")
data.frame(file = written,
           exists = file.exists(file.path(fpath, written)),
           row.names = NULL)
```

# Reproducibility

```{r}
#| label: session-info
#| echo: true

if (requireNamespace("sessioninfo", quietly = TRUE)) {
  sessioninfo::session_info()
} else {
  print(sessionInfo())
}

if (requireNamespace("renv", quietly = TRUE)) {
  try(print(renv::status()), silent = TRUE)
}
```
