---
title: "Getting Started with ratioPCR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ratioPCR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

The `ratioPCR` package is designed to read, process, and analyze quantitative real-time PCR data from ThermoFisher instruments using a clean and intuitive S3 object system.

```{r setup}
library(ratioPCR)
```

## 1. Importing Data

The starting point of the workflow is the `rpcr_import()` function. It reads the raw export file and automatically passes it to `rpcr_wrap()`. The result is a specialized data object of the S3 class `rpcr_raw`.

*(Note: If you leave the filepath empty, a file selection dialog will open automatically).*

```{r import, eval = FALSE}
# Interactive file import
pcr_data <- rpcr_import()
```

For this tutorial, we simulate the output generated after the import:

```{r dummy_data}
# Building a dummy data frame simulating the rpcr_raw structure
dummy_df <- data.frame(
  well = rep(c("A1", "A2"), each = 40),
  cycle = rep(1:40, 2),
  target = rep(c("Gene_A_incl", "Gene_A_excl"), each = 40),
  sample = "Sample_1",
  FA = runif(80, 0.1, 10),
  FB = runif(80, 0.1, 10),
  tPSI = 0.5, # Required for calibration
  stringsAsFactors = FALSE
)

# Convert the dummy data to the S3 class
pcr_data <- rpcr_wrap(dummy_df)
```

## 2. Inspecting the S3 Object

Thanks to the `rpcr_raw` S3 class, you do not need to memorize complex commands to see what is inside your data. Standard base R functions like `print()` adapt automatically to provide a clean overview.

```{r print}
print(pcr_data)
```

## 3. Visualization with ggplot2 and Patchwork

The core feature for quality control is the custom `plot()` method. It utilizes `rpcr_plotData` and `rpcr_plotBar`. If you choose `plot_type = "all"`, the generated charts are automatically combined into a single grid layout using the `patchwork` package.

```{r plot, warning = FALSE, message = FALSE}
# Plot all available diagnostic charts in a 2-column grid
plot(pcr_data, plot_type = "all", signal = "Rn")
```

You can also request specific chart types by changing the arguments:

```{r plot_single, warning = FALSE}
# Request only the bar charts
plot(pcr_data, plot_type = "cycle", signal = "Rn")
```
